---
title: "Corrected reanalysis of Lam et al. (2022)"
subtitle: "Fresh-leaf C:N and other corrections, compared step by step with the published analysis"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Version-pinned piecewiseSEM 2.1.2 lives in the project-local library
# .Rlib/ (gitignored); see the README.
.libPaths(c(normalizePath(".Rlib"), .libPaths()))
stopifnot(packageVersion("piecewiseSEM") == "2.1.2")

library(tidyverse)
library(knitr)
library(kableExtra)
library(lme4)
library(piecewiseSEM)
```

# Purpose

`historical_reproduction.Rmd` reproduces the published analysis of Lam et al. (2022, *Ecosystems* 25: 1006–1019) as it was run. This notebook corrects it and measures what each correction changes. The main correction is that the species' leaf C:N ratios, which the Methods describe as measured on fresh leaves, mixed fresh and senesced leaves for 13 of the 25 analysis species (one of them, *Pentace triptera*, had senesced leaves only).

The corrections are applied one step at a time, so that each effect can be reported on its own:

```{r scenarios-table}
scenarioInfo <- tribble(
  ~scenario, ~description,
  "A",   "Historical analysis, as published (must reproduce reference/).",
  "A24", "A on the 24 species with a fresh-leaf C:N, i.e. without Pentace triptera (PTR); isolates the effect of the smaller species set.",
  "B1-ovr", "A24 with fresh-leaf C:N only, but keeping the historical CTE override (CTE enters the models with the species mean C:N); separates the effect of the fresh-leaf values from that of removing the override.",
  "B1",  "A24 with fresh-leaf C:N only (mean of per-sample ratios), no CTE override, and the duplicated SEM (model 55) fixed.",
  "B2",  "B1 plus the Madhuca fix: the Q6 'Madhuca sp.' trees are the same taxon, which restores the Q6 Madhuca row. The final corrected analysis.",
  "B2-CTE", "Sensitivity: B2 with the C:N of Calophyllum tetrapterum (CTE) set to NA, because its two samples are only inferred to be fresh."
)
scenarioInfo |>
  kable(col.names = c("Scenario", "Description")) |>
  kable_styling(full_width = FALSE)
```

Decisions behind these steps (recorded in the working notes, D1–D5 and D11):

- **PTR (D1).** *Pentace triptera* has no fresh-leaf C:N, so its fresh C:N is `NA`. It drops from the C:N analyses and from the PCA, which needs complete cases.
- **Common data set (D11).** AICc can only be compared among models fitted to the same data, and the 108 SEMs are ranked together. So every model in B is fitted to the complete-case species set: PTR's two species–plot rows drop from all of them, not only from the C:N and PC models. This is provisional; imputing PTR's C:N is an alternative to consider later.
- **CTE (D2).** The two *Calophyllum tetrapterum* samples sit in the fresh-leaf block but have senesced-like N; their status is "inferred fresh, pending verification". B uses them as fresh and drops the historical override; B2-CTE treats them as missing.
- **Aggregation (D3).** As historically, species C:N is the mean of the per-sample ratios; only the rows used change.
- ***Madhuca* (D4).** All *Madhuca* in the plots, including the three Q6 trees named "Madhuca sp.", are one taxon. It is applied as a separate step (B2).
- **Model 55 (D5).** The plot-type SEM for hypothesis i with SLA pairs the canopy model with SIV → SLA instead of SLA → SIV. B uses the intended model.

Unchanged from the published analysis: total canopy production and its *t*-tests (Fig. 2), and the phenology GAMM (Fig. 3). Neither uses C:N or species-level tree names. The square-root transformation of SIV, which the paper does not mention, and the inverse-plot-number weights are kept throughout, because they are part of the published model rather than errors.

# Leaf C:N: historical versus fresh-leaf values

```{r cn}
chns <- bind_rows(
  mutate(read.csv("data/CHNS v3_senesced.csv"), leaf = "senesced"),
  mutate(read.csv("data/CHNS v3_fresh.csv"), leaf = "fresh"))

# Historical (all rows) and fresh-leaf species means of per-sample C:N
cnBySpecies <- chns |>
  summarise(CN_hist = mean(Ratio),
            CN_fresh = if (any(leaf == "fresh")) {
              mean(Ratio[leaf == "fresh"])
            } else NA_real_,
            n_fresh = sum(leaf == "fresh"),
            n_senesced = sum(leaf == "senesced"),
            .by = Species)
```

```{r cn-table}
# Analysis species (published set), from the historical reference output
speciesHist <- read_csv("reference/historical_species_traits_pca.csv",
                        show_col_types = FALSE)
sppPlotHist <- read_csv("reference/historical_species_plot.csv",
                        show_col_types = FALSE)
cnCompare <- sppPlotHist |>
  distinct(species, code) |>
  filter(code %in% speciesHist$code) |>
  left_join(cnBySpecies, by = c(species = "Species")) |>
  mutate(samples = case_when(
           n_fresh == 0 ~ "senesced only",
           n_senesced == 0 ~ "fresh only",
           .default = "fresh + senesced"),
         difference = CN_fresh - CN_hist) |>
  arrange(samples, difference)

cnCompare |>
  select(species, code, n_fresh, n_senesced, samples, CN_hist, CN_fresh,
         difference) |>
  kable(digits = 2,
        col.names = c("Species", "Code", "Fresh samples",
                      "Senesced samples", "Samples used historically",
                      "C:N, historical", "C:N, fresh only",
                      "Difference")) |>
  kable_styling(font_size = 12) |>
  scroll_box(height = "400px")
```

Of the 25 analysis species, `r sum(cnCompare$samples == "fresh + senesced")` had their C:N averaged over fresh and senesced leaves, `r sum(cnCompare$samples == "senesced only")` (*Pentace triptera*) over senesced leaves only, and `r sum(cnCompare$samples == "fresh only")` over fresh leaves only. Among the mixed species, the fresh-leaf value is lower by `r round(-max(cnCompare$difference[cnCompare$samples == "fresh + senesced"]), 2)` to `r round(-min(cnCompare$difference[cnCompare$samples == "fresh + senesced"]), 2)`.

```{r fig-cn, fig.width = 6, fig.height = 5, fig.cap = "Species C:N used in the published analysis against fresh-leaf C:N. The line is 1:1. PTR has no fresh-leaf value."}
cnCompare |>
  filter(!is.na(CN_fresh)) |>
  ggplot(aes(CN_fresh, CN_hist, colour = samples, label = code)) +
  geom_abline(linetype = 2) +
  geom_text(size = 3.5) +
  scale_colour_manual(values = c("fresh + senesced" = "#B2182B",
                                 "fresh only" = "grey40")) +
  labs(x = "C:N, fresh leaves only", y = "C:N, as published",
       colour = "Samples used\nhistorically") +
  coord_equal() +
  theme_bw()
```

# Species–plot data

The species–plot table is rebuilt as in `historical_reproduction.Rmd` (see there for the step-by-step explanation), with one switch for the *Madhuca* fix.

```{r spp-plot}
lamina <- read.csv("data/Cleaned lamina+PJ.csv") |>
  mutate(date1 = as.Date(General.date, format = "%d/%m/%Y"))
treeRaw <- read.csv("data/ten plot trees.csv") |>
  na.omit() |>
  filter(species != "") |>
  mutate(ba = pi * (dbh_2018 / 2)^2)
ssi <- read.csv("../ecophysio_traits/raw_data/SSI Jan21.csv")

sppSel <- c(
  "Adenanthera malayana", "Alstonia pneumatophora", "Aporosa falcifera",
  "Aporosa frutescens", "Aporosa prainiana", "Archidendron clypearia",
  "Artocarpus kemando", "Bhesa paniculata", "Buchanania sessifolia",
  "Calophyllum tetrapterum var. tetrapterum",
  "Calophyllum wallichianum var. incrassatum", "Campnosperma squamatum",
  "Canarium pilosum", "Diospyros maingayi", "Dysoxylum cauliflorum",
  "Elaeocarpus mastersii", "Elaeocarpus stipularis", "Gironniera nervosa",
  "Gluta wallichii", "Gynotroches axillaris", "Hevea brasiliensis",
  "Knema malayana", "Lindera lucida", "Lithocarpus ewyckii",
  "Litsea firma", "Lophopetalum multinervium", "Maasia glauca",
  "Madhuca tomentosa", "Mallotus paniculatus", "Myristica cinnamomea",
  "Nothaphoebe umbelliflora", "Pentace triptera", "Pometia pinnata",
  "Prunus polystachya", "Pternandra echinata", "Rhodamnia cinerea",
  "Sterculia cordata", "Sterculia macrophylla", "Timonius wallichianus",
  "Vatica pauciflora ", "Xanthophyllum flavescens")

# Species-plot litter production, BA and SIV (C:N is added per scenario).
# madhucaFix: treat the Q6 "Madhuca sp." trees as "Madhuca tomentosa", the
# name used in the litter, CHNS and trait data. (Which name to print is
# cosmetic; the litter name is kept so that the joins work.)
buildSppPlot <- function(madhucaFix) {
  tree <- treeRaw
  if (madhucaFix) {
    tree$species[tree$species == "Madhuca sp."] <- "Madhuca tomentosa"
  }
  lastDate <- max(lamina$date1, na.rm = TRUE)
  duration <- lamina |>
    filter(species %in% sppSel) |>
    summarise(start = min(date1, na.rm = TRUE),
              end = max(date1, na.rm = TRUE), .by = species) |>
    mutate(duration = as.numeric(if_else(end == start, lastDate - start,
                                         end - start)))
  madhucaSsi <- filter(ssi, X == "Madhuca sp.")
  expand_grid(plot = sort(unique(lamina$Plot)),
              species = intersect(sppSel, tree$species)) |>
    left_join(summarise(lamina, litter.collection = sum(Dry.Mass),
                        .by = c(Plot, species)),
              by = c(plot = "Plot", "species")) |>
    left_join(summarise(tree, ten.ba = sum(ba), .by = c(plot, species)),
              by = c("plot", "species")) |>
    replace_na(list(litter.collection = 0, ten.ba = 0)) |>
    left_join(select(duration, species, duration), by = "species") |>
    mutate(litter.production = litter.collection / 4 / (duration / 365),
           code = toupper(paste0(substr(word(species, 1), 1, 1),
                                 substr(word(species, 2), 1, 2)))) |>
    filter(litter.production != 0) |>
    filter(ten.ba != 0) |>
    # "Madhuca tomentosa" takes the SSI file's "Madhuca sp." values
    left_join(select(ssi, X, SSI = ssi.ba, iv.wet), by = c(species = "X")) |>
    mutate(SSI = replace_na(SSI, madhucaSsi$ssi.ba),
           iv.wet = replace_na(iv.wet, madhucaSsi$iv.wet))
}
sppPlot <- list(asPublished = buildSppPlot(FALSE),
                madhucaFixed = buildSppPlot(TRUE))

# The Madhuca fix should add exactly one row: Q6 Madhuca
madhucaNew <- anti_join(sppPlot$madhucaFixed, sppPlot$asPublished,
                        by = c("plot", "species"))
stopifnot(nrow(madhucaNew) == 1, madhucaNew$plot == "Q6")
```

The *Madhuca* fix adds one species–plot row, *Madhuca* in Q6, with `r round(madhucaNew$litter.collection)` g of leaf litter and `r format(round(madhucaNew$ten.ba), big.mark = ",")` cm^2^ of BA from the three "Madhuca sp." trees. Its SIV is the SSI file's "Madhuca sp." value, as for the other *Madhuca* rows; whether that value already pools all *Madhuca* trees in the 40 plots it was computed from could not be checked here, because those data are not in this repository.

# The analysis pipeline

One function runs the path analysis for a scenario: PCA, component models, the 108 SEMs and model selection. Its steps are exactly those of `historical_reproduction.Rmd`, with switches for the corrections.

```{r traits}
# Species trait means (unchanged by any correction)
traitsHj <- read.csv("data/HJ leaf data full.csv") |>
  mutate(across(c(thickness, ldmc, SLA),
                \(x) suppressWarnings(as.numeric(x))),
         code = toupper(substr(indiv, 1, 3))) |>
  summarise(SLA = mean(SLA, na.rm = TRUE),
            LDMC = mean(ldmc, na.rm = TRUE) / 1000,
            LT = mean(thickness, na.rm = TRUE),
            .by = code)
traitsPj <- read.csv("data/PJ leaf dat.csv") |>
  mutate(meanlt = rowMeans(across(starts_with("MESOPHYLL_THICKNESS")),
                           na.rm = TRUE)) |>
  summarise(SLA = mean(SLA_cm2.g.1, na.rm = TRUE),
            LDMC = mean(LDMC, na.rm = TRUE),
            LT = mean(meanlt),
            .by = SPECIES) |>
  rename(code = SPECIES)

# Historical PCA loadings, to align the arbitrary signs of the PCs
loadingsHist <- read_csv("reference/historical_pca_loadings.csv",
                         show_col_types = FALSE)
```

```{r pipeline}
traitNames <- c("SLA", "LT", "LDMC", "CNratio", "PC1", "PC2")
plotTypes <- c(noPlot = FALSE, plot = TRUE)

runScenario <- function(sppPlot, cn, cteOverride, fixModel55) {
  # C:N per species; species without a value drop from everything (D11)
  sppPlot <- sppPlot |>
    left_join(cn, by = "species") |>
    filter(!is.na(CNratio))
  traitsOnly <- bind_rows(traitsHj, traitsPj) |>
    inner_join(summarise(sppPlot, CNratio = mean(CNratio), .by = code),
               by = "code") |>
    arrange(code)

  pca <- prcomp(select(traitsOnly, SLA, LDMC, LT, CNratio), scale. = TRUE)
  # Align each PC's sign with the historical PC (signs are arbitrary)
  flip <- sign(colSums(pca$rotation[, 1:2] *
                         as.matrix(loadingsHist[match(rownames(pca$rotation),
                                                      loadingsHist$trait),
                                                c("PC1", "PC2")])))
  pca$rotation[, 1:2] <- sweep(pca$rotation[, 1:2], 2, flip, `*`)
  pca$x[, 1:2] <- sweep(pca$x[, 1:2], 2, flip, `*`)

  # The historical merge() chain fixes the row order, to which the lmer
  # refits in the d-separation tests are slightly sensitive
  mergeHist <- function(x, trait) {
    rbind(merge(x, as.data.frame(traitsHj[c("code", trait)]), by = "code"),
          merge(x, as.data.frame(traitsPj[c("code", trait)]), by = "code"))
  }
  litter <- sppPlot |>
    arrange(species, plot) |>
    as.data.frame() |>
    mergeHist("SLA") |>
    mergeHist("LT") |>
    mergeHist("LDMC") |>
    merge(data.frame(code = traitsOnly$code, pca$x), by = "code") |>
    as_tibble()
  if (cteOverride) {
    litter$CNratio[litter$code == "CTE"] <- mean(traitsOnly$CNratio)
  }
  litter <- litter |>
    mutate(loglitter = log(litter.production), logba = log(ten.ba),
           weights = 1 / n(), .by = species) |>
    mutate(plotwet = as.numeric(plot %in% c("Q3", "Q4", "Q6", "Q9", "Q10")),
           SLA = SLA / 100,
           iv.wet = sqrt(iv.wet)) |>
    as.data.frame()

  # Component models. The data are spliced into each call so that
  # piecewiseSEM can refit the models for its d-separation tests.
  fitLm <- function(f) {
    eval(bquote(lm(.(f), weights = weights, data = .(litter))))
  }
  fitLmer <- function(fixed, plotType) {
    f <- reformulate(c("logba", fixed, if (plotType) "plotwet",
                       "(1 | plot)", "(logba | species)"),
                     response = "loglitter")
    eval(bquote(lmer(.(f), data = .(litter))))
  }
  modS0 <- map(set_names(traitNames), \(t) fitLm(reformulate("1", t)))
  modS <- map(set_names(traitNames), \(t) fitLm(reformulate("iv.wet", t)))
  modD <- map(set_names(traitNames), \(t) fitLm(reformulate(t, "iv.wet")))
  modD0 <- fitLm(iv.wet ~ 1)
  modL0 <- map(plotTypes, \(p) fitLmer(NULL, p))
  modL_siv <- map(plotTypes, \(p) fitLmer("iv.wet", p))
  modL_trait <- map(plotTypes, \(p)
    map(set_names(traitNames), \(t) fitLmer(t, p)))
  modL_sivTrait <- map(plotTypes, \(p)
    map(set_names(traitNames), \(t) fitLmer(c("iv.wet", t), p)))

  # The 108 SEMs, in the historical order
  sems <- map(names(plotTypes), \(p) {
    s <- list(
      i    = map(traitNames, \(t) psem(modL_sivTrait[[p]][[t]], modD[[t]])),
      ii   = map(traitNames, \(t) psem(modL_sivTrait[[p]][[t]], modS[[t]])),
      vi   = map(traitNames, \(t) psem(modD[[t]], modL_siv[[p]])),
      vii  = map(traitNames, \(t) psem(modS[[t]], modL_trait[[p]][[t]])),
      iii  = map(traitNames, \(t)
        psem(modL_sivTrait[[p]][[t]], modS0[[t]], modD0)),
      iv   = map(traitNames, \(t) psem(modL_siv[[p]], modD0, modS0[[t]])),
      viii = map(traitNames, \(t) psem(modD[[t]], modS0[[t]], modL0[[p]])),
      ix   = map(traitNames, \(t) psem(modS[[t]], modD0, modL0[[p]])),
      v    = map(traitNames, \(t)
        psem(modL_trait[[p]][[t]], modS0[[t]], modD0)))
    tibble(plotType = p, BH = rep(names(s), each = length(traitNames)),
           trait = rep(traitNames, length(s)),
           sem = unname(list_flatten(s)))
  }) |>
    list_rbind() |>
    mutate(id = row_number(), .before = 1)
  if (!fixModel55) {
    # Historical bug (D5): model 55 duplicates the BH ii model 61
    sems$sem[[55]] <- psem(modL_sivTrait$plot$SLA, modS$SLA)
  }

  # Model selection, with the historical rounding of weights
  sems <- sems |>
    mutate(AICc = map_dbl(sem, \(m)
             tryCatch(AIC(m, aicc = TRUE), error = \(e) NA_real_)),
           dAICc = AICc - min(AICc, na.rm = TRUE),
           RL = round(exp(-0.5 * dAICc), 5),
           weight_sow = round(RL / sum(RL, na.rm = TRUE), 3)) |>
    arrange(dAICc) |>
    mutate(rank = row_number(),
           dAICc_tab = round(dAICc, 2),
           weight = round(round(exp(-0.5 * dAICc_tab), 5) /
                            sum(round(exp(-0.5 * dAICc_tab), 5),
                                na.rm = TRUE), 3),
           R2 = map_chr(sem, \(m) tryCatch(
             paste(round(100 * rsquared(m)$Marginal, 2), collapse = " / "),
             error = \(e) NA_character_)),
           FisherC = map_dbl(sem, \(m) tryCatch(
             fisherC(dSep(m, .progressBar = FALSE))$Fisher.C,
             error = \(e) NA_real_)),
           formula = map_chr(sem, \(m) {
             comps <- m[map_lgl(m, \(x) inherits(x, c("lm", "merMod")))]
             map_chr(comps, \(x) deparse1(formula(x))) |>
               str_remove(fixed(" + (1 | plot) + (logba | species)")) |>
               paste(collapse = "; ")
           }))

  list(litter = litter, traitsOnly = traitsOnly, pca = pca,
       modD = modD, modS = modS, modL_siv = modL_siv,
       modL_trait = modL_trait, modL_sivTrait = modL_sivTrait,
       modL0 = modL0, sems = sems,
       sow = summarise(sems, SoW = sum(weight_sow), .by = BH))
}
```

# Running the scenarios

```{r run-scenarios, cache = TRUE}
# CACHED (about 20 min to run). knitr cannot track the uncached chunks this
# depends on, so delete corrected_reanalysis_cache/ after changing any chunk
# above this one.
cnHist <- select(cnBySpecies, species = Species, CNratio = CN_hist)
cnFresh <- select(cnBySpecies, species = Species, CNratio = CN_fresh)
cteName <- "Calophyllum tetrapterum var. tetrapterum"
ptrName <- "Pentace triptera"

results <- list(
  A = runScenario(sppPlot$asPublished, cnHist,
                  cteOverride = TRUE, fixModel55 = FALSE),
  A24 = runScenario(sppPlot$asPublished,
                    mutate(cnHist, CNratio = if_else(species == ptrName,
                                                     NA, CNratio)),
                    cteOverride = TRUE, fixModel55 = FALSE),
  `B1-ovr` = runScenario(sppPlot$asPublished, cnFresh,
                         cteOverride = TRUE, fixModel55 = FALSE),
  B1 = runScenario(sppPlot$asPublished, cnFresh,
                   cteOverride = FALSE, fixModel55 = TRUE),
  B2 = runScenario(sppPlot$madhucaFixed, cnFresh,
                   cteOverride = FALSE, fixModel55 = TRUE),
  `B2-CTE` = runScenario(sppPlot$madhucaFixed,
                         mutate(cnFresh, CNratio = if_else(species == cteName,
                                                           NA, CNratio)),
                         cteOverride = FALSE, fixModel55 = TRUE))
```

```{r check-a}
# Scenario A must reproduce the historical reference outputs
semHist <- read_csv("reference/historical_sem_table.csv",
                    show_col_types = FALSE)
aCheck <- results$A$sems |>
  select(id, AICc, rank, weight) |>
  inner_join(select(semHist, id, AICc, rank, weight), by = "id",
             suffix = c("", "_ref"))
# AIC.psem() rounds AICc to 3 dp, and the d-separation refits are
# sensitive in that last digit to how the data reach the model call, so a
# difference of a few thousandths is tolerated; ranks must be identical.
aMaxDiff <- max(abs(aCheck$AICc - aCheck$AICc_ref))
stopifnot(nrow(aCheck) == 108, aMaxDiff <= 0.005,
          all(aCheck$rank == aCheck$rank_ref),
          all(aCheck$weight == aCheck$weight_ref))
```

Scenario A reproduces the historical analysis: all 108 SEMs have the same rank and Akaike weight as in `reference/historical_sem_table.csv`, and their AICc values differ by at most `r aMaxDiff`, in the third decimal, where `AIC.psem()` rounds.

```{r sample-sizes}
scenarioOrder <- names(results)
imap(results, \(r, s) tibble(
  scenario = s,
  species = n_distinct(r$litter$species),
  rows = nrow(r$litter),
  CN_CTE = if (any(r$litter$code == "CTE")) {
    unique(r$litter$CNratio[r$litter$code == "CTE"])
  } else NA_real_,
  PC1 = 100 * summary(r$pca)$importance[2, "PC1"],
  PC2 = 100 * summary(r$pca)$importance[2, "PC2"],
  semsWithAICc = sum(!is.na(r$sems$AICc)))) |>
  list_rbind() |>
  kable(digits = 2,
        col.names = c("Scenario", "Species", "Species–plot rows",
                      "C:N of CTE in models", "PC1 (%)", "PC2 (%)",
                      "SEMs with AICc"),
        caption = "Data and PCA by scenario.") |>
  kable_styling(full_width = FALSE)
```

# What changes

## PCA

```{r pca-loadings}
imap(results, \(r, s)
  as_tibble(r$pca$rotation[, 1:2], rownames = "trait") |>
    mutate(scenario = s)) |>
  list_rbind() |>
  pivot_wider(names_from = scenario, values_from = c(PC1, PC2),
              names_glue = "{.value} {scenario}") |>
  kable(digits = 2,
        caption = "PCA loadings by scenario (signs aligned with A).") |>
  kable_styling(font_size = 12)
```

## Trait–SIV and canopy-production paths

The published Results quote these path coefficients (95% Wald CIs, estimate ± 1.96 SE, as in the paper).

```{r component-paths}
coefRow <- function(mod, term) {
  cf <- coef(summary(mod))[term, ]
  est <- cf[["Estimate"]]
  se <- cf[["Std. Error"]]
  tibble(estimate = est, lwr = est - 1.96 * se, upr = est + 1.96 * se,
         R2 = if (inherits(mod, "lm")) summary(mod)$r.squared else NA)
}
paths <- imap(results, \(r, s) bind_rows(
  map(set_names(traitNames), \(t) coefRow(r$modD[[t]], t)) |>
    list_rbind(names_to = "trait") |>
    mutate(path = paste(trait, "-> SIV")),
  coefRow(r$modS$LDMC, "iv.wet") |> mutate(path = "SIV -> LDMC"),
  coefRow(r$modL_siv$noPlot, "iv.wet") |> mutate(path = "SIV -> canopy"),
  coefRow(r$modL_siv$noPlot, "logba") |> mutate(path = "BA -> canopy")) |>
  mutate(scenario = s)) |>
  list_rbind() |>
  select(-trait)

paths |>
  mutate(value = sprintf("%.3f [%.3f, %.3f]%s", estimate, lwr, upr,
                         if_else(is.na(R2), "",
                                 sprintf("; R² %.1f%%", 100 * R2)))) |>
  select(path, scenario, value) |>
  pivot_wider(names_from = scenario, values_from = value) |>
  kable(caption = "Path coefficients [95% CI] by scenario.") |>
  kable_styling(font_size = 11)
```

```{r fig-cn-siv, fig.width = 8, fig.height = 4.5, fig.cap = "C:N → SIV coefficient (95% CI) by scenario. The published value is 0.009 [0.004, 0.013]."}
paths |>
  filter(path %in% c("CNratio -> SIV", "PC1 -> SIV")) |>
  mutate(scenario = factor(scenario, levels = scenarioOrder)) |>
  ggplot(aes(scenario, estimate, ymin = lwr, ymax = upr)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_pointrange() +
  facet_wrap(~ path, scales = "free_y") +
  labs(x = "Scenario", y = "Coefficient (95% CI)") +
  theme_bw()
```

## Model selection

```{r top-models}
topModels <- imap(results, \(r, s)
  r$sems |>
    filter(rank <= 10) |>
    transmute(scenario = s, rank, id, BH, trait, plotType,
              AICc = round(AICc, 2), dAICc = dAICc_tab, weight)) |>
  list_rbind()

topModels |>
  mutate(model = sprintf("%s %s%s (%.3f)", BH, trait,
                         if_else(plotType == "plot", " +plot", ""),
                         weight)) |>
  select(rank, scenario, model) |>
  pivot_wider(names_from = scenario, values_from = model) |>
  kable(caption = "Top ten SEMs by scenario: hypothesis, trait (Akaike weight).") |>
  kable_styling(font_size = 11)
```

```{r published-top-ranks}
# Where do the published top ten models rank in each scenario?
publishedTop <- results$A$sems |>
  filter(rank <= 10) |>
  select(id, BH, trait, plotType, rank_A = rank)
imap(results[-1], \(r, s)
  select(r$sems, id, rank, weight) |>
    rename_with(\(x) paste0(x, "_", s), -id)) |>
  reduce(left_join, by = "id", .init = publishedTop) |>
  kable(caption = "Ranks (and weights) of the ten published top models.") |>
  kable_styling(font_size = 11)
```

```{r sow}
sowAll <- imap(results, \(r, s) mutate(r$sow, scenario = s)) |>
  list_rbind()
sowAll |>
  mutate(SoW = round(100 * SoW, 1)) |>
  pivot_wider(names_from = scenario, values_from = SoW) |>
  arrange(desc(A)) |>
  kable(caption = "Summed Akaike weights (%) per biological hypothesis.") |>
  kable_styling(full_width = FALSE)
```

```{r fig-sow, fig.width = 8, fig.height = 5, fig.cap = "Summed Akaike weights per biological hypothesis, by scenario (Fig. 5 of the paper is scenario A)."}
sowAll |>
  mutate(scenario = factor(scenario, levels = scenarioOrder),
         BH = fct_reorder(BH, SoW, .fun = max)) |>
  ggplot(aes(100 * SoW, BH, fill = scenario)) +
  geom_col(position = position_dodge(0.85), width = 0.8) +
  scale_fill_manual(values = c("grey70", "grey45", "#D1E5F0", "#67A9CF",
                               "#2166AC", "#B2182B")) +
  labs(x = "Summed Akaike weight (%)", y = "Biological hypothesis",
       fill = "Scenario") +
  theme_bw()
```

## The C:N models specifically

The Abstract and Discussion state that a high leaf C:N ratio is associated with adaptation to the swamp. The SEMs that carry this are the C:N models of each hypothesis:

```{r cn-models}
imap(results, \(r, s)
  r$sems |>
    filter(trait == "CNratio", plotType == "noPlot") |>
    transmute(BH, scenario = s,
              value = sprintf("%d (%.3f)", rank, weight))) |>
  list_rbind() |>
  pivot_wider(names_from = scenario, values_from = value) |>
  kable(caption = "Rank (Akaike weight) of each C:N SEM without plot type.") |>
  kable_styling(full_width = FALSE)
```

## The influence of CTE

```{r fig-cte, fig.width = 9, fig.height = 4.5, fig.cap = "SIV (square-rooted) against species C:N, as published (A) and with fresh-leaf C:N (B2). CTE is in red; in A it enters the models with the overridden value. Lines are unweighted least-squares fits, for orientation only."}
bind_rows(
  results$A$litter |> distinct(code, CNratio, iv.wet) |> mutate(scenario = "A"),
  results$B2$litter |> distinct(code, CNratio, iv.wet) |>
    mutate(scenario = "B2")) |>
  ggplot(aes(CNratio, iv.wet, label = code)) +
  geom_smooth(method = "lm", se = FALSE, colour = "grey60") +
  geom_text(aes(colour = code == "CTE"), size = 3.2, show.legend = FALSE) +
  scale_colour_manual(values = c("black", "#B2182B")) +
  facet_wrap(~ scenario) +
  labs(x = "Leaf C:N used in the models", y = "sqrt(SIV)") +
  theme_bw()
```

# Findings

```{r findings-helpers}
pathText <- function(s, p) {
  x <- filter(paths, scenario == s, path == p)
  sprintf("%.3f [%.3f, %.3f], R² %.1f%%", x$estimate, x$lwr, x$upr,
          100 * x$R2)
}
sowText <- function(s, h) {
  sprintf("%.1f%%", 100 * sowAll$SoW[sowAll$scenario == s & sowAll$BH == h])
}
topText <- function(s) {
  x <- filter(results[[s]]$sems, rank == 1)
  sprintf("BH %s with %s (weight %.3f)", x$BH, x$trait, x$weight)
}
rankCn <- function(s) {
  with(results[[s]]$sems, rank[BH == "vi" & trait == "CNratio" &
                                 plotType == "noPlot"])
}
```

- **C:N → SIV.** Published (A): `r pathText("A", "CNratio -> SIV")`. Dropping PTR alone (A24) strengthens it: `r pathText("A24", "CNratio -> SIV")`. With fresh-leaf C:N and the historical CTE override (B1-ovr): `r pathText("B1-ovr", "CNratio -> SIV")`. With fresh-leaf C:N and CTE's own value (B1): `r pathText("B1", "CNratio -> SIV")`; after the *Madhuca* fix (B2): `r pathText("B2", "CNratio -> SIV")`. Without CTE (B2-CTE): `r pathText("B2-CTE", "CNratio -> SIV")`. The strength of the C:N–habitat association therefore depends mainly on how CTE is treated, not on the fresh/senesced mixing as such.
- **The C:N chain model (BH vi, C:N → SIV → canopy)** ranks `r rankCn("A")` in A, `r rankCn("B1-ovr")` in B1-ovr, `r rankCn("B2")` in B2 and `r rankCn("B2-CTE")` in B2-CTE.
- **Top model.** A: `r topText("A")`; B2: `r topText("B2")`; B2-CTE: `r topText("B2-CTE")`.
- **Summed weights.** BH vi (traits → SIV → canopy) has `r sowText("A", "vi")` in A and `r sowText("B2", "vi")` in B2; BH i has `r sowText("A", "i")` and `r sowText("B2", "i")`; BH vii has `r sowText("A", "vii")` and `r sowText("B2", "vii")`. BH vi remains the best-supported hypothesis in every scenario.
- **Other paths.** SLA → SIV (`r pathText("B2", "SLA -> SIV")` in B2) and PC1 → SIV (`r pathText("B2", "PC1 -> SIV")`) hardly change; SIV → canopy remains non-significant (B2: `r with(filter(paths, scenario == "B2", path == "SIV -> canopy"), sprintf("%.2f [%.2f, %.2f]", estimate, lwr, upr))`).

# The corrected analysis (B2)

```{r b2-table1}
results$B2$sems |>
  filter(rank <= 10) |>
  select(rank, BH, formula, R2, FisherC, AICc, dAICc = dAICc_tab, weight) |>
  kable(digits = 3,
        caption = "Table 1, corrected (B2): the ten top-ranked SEMs.") |>
  kable_styling(font_size = 12)
```

```{r fig4-b2, fig.width = 7, fig.height = 11, fig.cap = "Figure 4, corrected (B2): a) PCA of leaf traits with fresh-leaf C:N (24 species); b) SIV (square-rooted) against SLA."}
r <- results$B2
pcaVar <- 100 * summary(r$pca)$importance[2, ]
pcaArrows <- as_tibble(r$pca$rotation, rownames = "trait") |>
  mutate(trait = recode(trait, CNratio = "C:N"))
fig4a <- bind_cols(code = r$traitsOnly$code, as_tibble(r$pca$x)) |>
  ggplot(aes(PC1, PC2)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_text(aes(label = code)) +
  geom_segment(data = pcaArrows,
               aes(x = 0, y = 0, xend = 3 * PC1, yend = 3 * PC2),
               colour = "forestgreen", linewidth = 1,
               arrow = arrow(length = unit(0.2, "cm"))) +
  geom_text(data = pcaArrows, aes(4 * PC1, 3.5 * PC2, label = trait),
            colour = "forestgreen", size = 5) +
  labs(x = sprintf("PC 1 (%.1f%% of variance)", pcaVar[["PC1"]]),
       y = sprintf("PC 2 (%.1f%% of variance)", pcaVar[["PC2"]]),
       title = "a)") +
  theme_bw()
fig4b <- r$litter |>
  distinct(code, SLA, iv.wet) |>
  ggplot(aes(SLA * 100, iv.wet, label = code)) +
  geom_text() +
  labs(x = expression(SLA~(cm^2~g^-1)),
       y = "Habitat specialisation (sqrt of swamp indicator value)",
       title = "b)") +
  theme_bw()
patchwork::wrap_plots(fig4a, fig4b, ncol = 1)
```

# Outputs

```{r outputs}
# Written for the comparison and write-up (Stage 4)
for (s in names(results)) {
  stem <- paste0("reference/corrected_", s)
  write_csv(select(results[[s]]$sems, -sem), paste0(stem, "_sem_table.csv"))
}
write_csv(paths, "reference/corrected_paths.csv")
write_csv(sowAll, "reference/corrected_sum_of_weights.csv")
write_csv(cnCompare, "reference/corrected_cn_by_species.csv")
write_csv(
  imap(results, \(r, s)
    bind_cols(scenario = s, r$traitsOnly, as_tibble(r$pca$x))) |>
    list_rbind(),
  "reference/corrected_species_traits_pca.csv")
```

The outputs are in `reference/` with the prefix `corrected_`, one SEM table per scenario plus the paths, summed weights, species C:N and species PCA scores.

# Session information

```{r session-info}
sessionInfo()
```
