---
title: "Lam et al. (2022): Results and Supplementary Material, reproduced"
subtitle: "Habitat adaptation mediates the influence of leaf traits on canopy productivity: evidence from a tropical freshwater swamp forest. *Ecosystems* 25: 1006–1019"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Version-pinned packages (piecewiseSEM 2.1.2, gamm4 0.2-6) live in the
# project-local library .Rlib/ (gitignored); see the README.
.libPaths(c(normalizePath(".Rlib"), .libPaths()))
stopifnot(packageVersion("piecewiseSEM") == "2.1.2")

library(tidyverse)
library(knitr)
library(kableExtra)

# Every comparison with a published (or historical) value is added to
# `checks`, which is shown in Note 3. A check passes when the reproduced
# value, rounded to the published precision, equals it.
checks <- tibble()
addCheck <- function(section, item, published, reproduced, digits) {
  checks <<- bind_rows(checks, tibble(
    section = section, item = item, published = published,
    reproduced = round(reproduced, digits),
    match = round(reproduced, digits) == published))
}

# Formatted values for the text, kept in `v` so that pasted paragraphs can
# cite them inline, e.g. `r v$prodMean`. Each paragraph's values are listed
# in a collapsible key under its paste slot.
v <- list()
valueKey <- function(key) {
  cat("<details><summary>Values for this paragraph</summary>\n\n")
  key |>
    mutate(inline = sprintf("<code>`r v$%s`</code>", name),
           reproduced = map_chr(name, \(n) v[[n]])) |>
    select(statistic, published, reproduced, inline) |>
    kable(escape = FALSE,
          col.names = c("Statistic", "Published", "Reproduced",
                        "Inline code")) |>
    kable_styling(full_width = FALSE, font_size = 12) |>
    cat()
  cat("\n\n</details>\n")
}
f <- function(x, digits = 2) sprintf(paste0("%.", digits, "f"), x)
```

```{css}
.slot {
  border: 2px dashed #B2182B; background: #FFF5F5; color: #8B1A1A;
  padding: 6px 10px; margin: 10px 0;
}
.preamble { background: #F4F6F8; padding: 10px 14px; border-radius: 4px; }
```

::: {.preamble}
**About this document (not part of the paper).** This notebook reproduces the Results and Supplementary Material of Lam et al. (2022) from the original input data, using the analysis as it was actually run (historical scripts, commit `74befc2`). Every number below is computed on knitting, and each figure and table is regenerated from the data. The paper's own text goes in the marked slots, pasted from the authors' manuscript; under each slot, the values it cites are listed with an inline expression to put in place of each number, so that the text updates on knitting. Material that is not in the paper (inputs, historical quirks, verification against the published values, software notes) is in the [Notes](#notes) at the end. The corrected analysis is in `corrected_reanalysis.Rmd`.

The reproduction keeps the original analysis exactly, including choices now known to be errors (Note 2), the most important being that leaf C:N ratios mixed fresh and senesced leaves. Almost all published numbers are reproduced exactly; the few that are not differ in the second decimal place (Note 3).
:::

# Results

## Total canopy production

```{r litter-data, include = FALSE}
lamina <- read.csv("data/Cleaned lamina+PJ.csv") |>
  mutate(date1 = as.Date(General.date, format = "%d/%m/%Y"),
         date2 = format(date1, "%Y-%m"))

# Rows with NA in any column are dropped (historical na.omit()), and so are
# trees without a species name.
tree <- read.csv("data/ten plot trees.csv") |>
  na.omit() |>
  filter(species != "") |>
  mutate(ba = pi * (dbh_2018 / 2)^2)

# The historical code picked these by factor-level index; the names below
# are what those indices resolve to (verified under R 4.3.3).
lianas <- c(
  "Agelaea borneensis", "Another climber", "Bauhinia semibifida",
  "Dalbergia pseudo-sissoo", "Derris amoena",
  "Fissistigma latifolium var. ovoideum", "Grewia laevigata",
  "Iodes ovalis", "Iodes velutina", "Leuconotis griffithii",
  "Smilax setosa", "Strychnos", "Uncaria cordata",
  "Uncaria cordata var. cordata f. sundaica", "Willughbeia coriacea")
```

```{r total-production, include = FALSE}
# Plot-level stand data for leaf litter of non-liana species (also the
# basis of the GAMM): litter per collection, plot BA, BA-normalised
# production.
plotBa <- summarise(tree, ba = sum(ba), .by = plot)
# aggregate() dropped rows with NA mass or date (historical formula method)
standLitter <- lamina |>
  filter(!species %in% lianas, !is.na(Dry.Mass), !is.na(date1)) |>
  summarise(Dry.Mass = sum(Dry.Mass), .by = c(date1, date2, Plot)) |>
  left_join(distinct(lamina, Plot, Condition), by = "Plot") |>
  left_join(plotBa, by = c(Plot = "plot")) |>
  mutate(prod = Dry.Mass / ba * 100^2 / 4)

# Per plot, in g m-2 yr-1 (the survey lasted about a year): all leaves,
# non-liana leaves, and BA-normalised non-liana leaves
plotProduction <- lamina |>
  summarise(all = sum(Dry.Mass) / 4, .by = Plot) |>
  left_join(summarise(standLitter,
                      bolian = sum(Dry.Mass) / 4,
                      bolianba = sum(prod),
                      .by = c(Plot, Condition)),
            by = "Plot") |>
  arrange(Plot)

seMean <- function(x) sd(x) / sqrt(length(x))
tTests <- c(all = "all", bolian = "bolian", bolianba = "bolianba") |>
  map(\(x) t.test(reformulate("Condition", x), data = plotProduction,
                  var.equal = TRUE))

addCheck("Fig. 2", "mean canopy production (g m-2 yr-1)", 768,
         mean(plotProduction$all), 0)
addCheck("Fig. 2", "SE of canopy production", 48,
         seMean(plotProduction$all), 0)
addCheck("Fig. 2", "total litter (kg)", 30.7,
         sum(lamina$Dry.Mass, na.rm = TRUE) / 1000, 1)
addCheck("Fig. 2", "t, all leaves", 1.33,
         abs(tTests$all$statistic), 2)
addCheck("Fig. 2", "P, all leaves", 0.221, tTests$all$p.value, 3)
addCheck("Fig. 2", "t, excluding lianas", 0.49,
         abs(tTests$bolian$statistic), 2)
addCheck("Fig. 2", "P, excluding lianas", 0.639, tTests$bolian$p.value, 3)
addCheck("Fig. 2", "t, BA-normalised", 6.81,
         abs(tTests$bolianba$statistic), 2)

v$litterKg <- f(sum(lamina$Dry.Mass, na.rm = TRUE) / 1000, 1)
v$nTraps <- "40"
v$prodMean <- f(mean(plotProduction$all), 0)
v$prodSe <- f(seMean(plotProduction$all), 0)
v$tAll <- f(abs(tTests$all$statistic))
v$dfT <- f(tTests$all$parameter, 0)
v$pAll <- f(tTests$all$p.value, 3)
v$tBolian <- f(abs(tTests$bolian$statistic))
v$pBolian <- f(tTests$bolian$p.value, 3)
v$tBa <- f(abs(tTests$bolianba$statistic))
v$pBa <- if (tTests$bolianba$p.value < 0.001) "< 0.001" else
  f(tTests$bolianba$p.value, 3)
```

::: {.slot}
**Paste here:** Results, *Total canopy production*, first paragraph (accepted manuscript l. 288–296).
:::

```{r key-p1, results = "asis"}
valueKey(tribble(
  ~statistic, ~published, ~name,
  "Leaf litter collected (kg)", "30.7", "litterKg",
  "Canopy production, mean (g m⁻² yr⁻¹)", "768", "prodMean",
  "Canopy production, SE", "48", "prodSe",
  "t, all leaves", "1.33", "tAll",
  "df", "8", "dfT",
  "P, all leaves", "0.221", "pAll",
  "t, liana-excluded", "0.49", "tBolian",
  "P, liana-excluded", "0.639", "pBolian",
  "t, per basal area", "6.81", "tBa",
  "P, per basal area", "< 0.001", "pBa"))
```

```{r weather, include = FALSE}
# Monthly NEA files, as read by the historical code. HISTORICAL QUIRK: an
# offset vector of months stops the reading at June 2019, although
# collections run to 30 July 2019. Rain at Upper Peirce (S69) is replaced
# by Mandai (S40) for March-June 2019, when S69 lacks most of its data.
dirNea <- "../../TFERP/LTFEM_aux/data/nea/weather.gov.sg/"
monthsRead <- format(seq(as.Date("2018-06-01"), as.Date("2019-06-01"),
                         by = "month"), "%Y%m")
readNea <- function(folder, station, months) {
  map(months, \(m)
    read.csv(file.path(dirNea, folder,
                       sprintf("DAILYDATA_%s_%s.csv", station, m)),
             fileEncoding = "CP1252", check.names = FALSE,
             colClasses = "character")[, 1:11]) |>
    list_rbind() |>
    set_names(c("station", "year", "month", "day", "rain", "r30", "r60",
                "r120", "tmean", "tmax", "tmin")) |>
    mutate(date1 = make_date(as.integer(year), as.integer(month),
                             as.integer(day)))
}
mandaiMonths <- c("201903", "201904", "201905", "201906")
temperature <- readNea("S122_khatib", "S122", monthsRead)
rain <- bind_rows(
  readNea("S69_upper_peirce_reservoir", "S69",
          setdiff(monthsRead, mandaiMonths)),
  readNea("S40_mandai", "S40", mandaiMonths))

# Missing values are coded with an em dash. HISTORICAL: missing rain is
# set to 0; a missing daily mean temperature takes that month's mean; a
# missing max or min stays NA.
weather <- temperature |>
  select(date1, tmean, tmax, tmin) |>
  left_join(select(rain, date1, rain), by = "date1") |>
  mutate(rain = as.numeric(if_else(rain == "—", "0", rain)),
         across(c(tmean, tmax, tmin), \(x) suppressWarnings(as.numeric(x))),
         ym = format(date1, "%Y%m")) |>
  mutate(tmean = if_else(is.na(tmean), mean(tmean, na.rm = TRUE), tmean),
         .by = ym)
```

```{r gamm-data, include = FALSE}
# For each collection date d: the 30 days before it (d-30 to d-1), and
# the 30 days before that (d-60 to d-31) for the lagged variables
windowMean <- function(x, d, from, to) {
  map_dbl(d, \(di) mean(x[weather$date1 >= di - from &
                            weather$date1 <= di - to]))
}
windowMax <- function(x, d) {
  map_dbl(d, \(di) max(x[weather$date1 >= di - 30 &
                           weather$date1 <= di - 1], na.rm = TRUE))
}
gammData <- standLitter |>
  mutate(rf = windowMean(weather$rain, date1, 30, 1),
         rf_lag = windowMean(weather$rain, date1, 60, 31),
         temp = windowMean(weather$tmean, date1, 30, 1),
         temp_lag = windowMean(weather$tmean, date1, 60, 31),
         temp_max = windowMax(weather$tmax, date1),
         temp_min = -windowMax(-weather$tmin, date1))

# Compare with the historical GAMM input file
gammHist <- read.csv("data/data for GAMM.csv") |>
  mutate(date1 = as.Date(date1))
gammCmp <- inner_join(gammData, gammHist, by = c("Plot", "date1"),
                      suffix = c("", "_hist"))
gammVars <- c("Dry.Mass", "ba", "prod", "rf", "rf_lag", "temp", "temp_lag",
              "temp_max", "temp_min")
gammMaxDiff <- map_dbl(set_names(gammVars), \(x)
  max(abs(gammCmp[[x]] - gammCmp[[paste0(x, "_hist")]])))
addCheck("GAMM data", "rows (data for GAMM.csv)", nrow(gammHist),
         nrow(gammData), 0)
addCheck("GAMM data", "rows matched", nrow(gammHist), nrow(gammCmp), 0)
addCheck("GAMM data", "max abs difference, all columns", 0,
         max(gammMaxDiff), 6)

# Collections whose 30-day window runs past the weather data
nTruncated <- sum(gammData$date1 - 1 > max(weather$date1))
```

```{r gamm-fit, include = FALSE}
library(mgcv)
library(gamm4)
library(MuMIn)
# The historical model selection, verbatim apart from the data name.
# CAUTION: the rows must be sorted by Plot (the random-effect grouping), as
# the historical merge() left them. Otherwise gamm4 0.2-6 returns the same
# estimates but wrong standard errors for the parametric terms.
tll <- gammData |>
  arrange(Plot, date1) |>
  as.data.frame()
uGammFull <- uGamm(log(prod) ~ Condition + s(as.numeric(date1))
  + rf + rf_lag + temp + temp_lag
  + temp_max + temp_min
  + Condition:rf + Condition:rf_lag + Condition:temp + Condition:temp_lag
  + Condition:temp_max + Condition:temp_min,
  random = ~(1|Plot), data = tll, lme4 = TRUE)
dredged <- dredge(uGammFull, m.lim = c(0, 4), rank = "AICc",
  subset = !((rf && rf_lag) || (temp && temp_lag) ||
    (temp_max && temp_min) || (temp && temp_max) ||
    (temp && temp_min) || (temp_lag && temp_max) || (temp_lag && temp_min)))

best <- gamm4(log(prod) ~ s(as.numeric(date1)) + Condition + temp,
              random = ~(1|Plot), data = tll)
bestCoefs <- summary(best$gam)$p.table
waldCi <- function(term) {
  bestCoefs[term, "Estimate"] +
    c(est = 0, lwr = -1.96, upr = 1.96) * bestCoefs[term, "Std. Error"]
}
# Dry relative to wet (the reference level is dry, D)
dryWet <- set_names(exp(-waldCi("ConditionW"))[c("est", "upr", "lwr")],
                    c("est", "lwr", "upr"))
tempEffect <- waldCi("temp")
r2Top <- r.squaredLR(get.models(dredged, subset = 1)[[1]])

addCheck("GAMM", "top-model weight > 0.90", 1,
         as.numeric(dredged$weight[1] > 0.9), 0)
addCheck("GAMM", "dAICc of 2nd model", 5.47, dredged$delta[2], 2)
addCheck("GAMM", "adj R² (LR) of top model (%)", 75,
         100 * attr(r2Top, "adj.r.squared"), 0)
addCheck("GAMM", "dry/wet multiplier", 2.01, dryWet[[1]], 2)
addCheck("GAMM", "dry/wet lower 95%", 1.57, dryWet[[2]], 2)
addCheck("GAMM", "dry/wet upper 95%", 2.57, dryWet[[3]], 2)
addCheck("GAMM", "temperature estimate", 0.40, tempEffect[["est"]], 2)
addCheck("GAMM", "temperature lower 95%", 0.27, tempEffect[["lwr"]], 2)
addCheck("GAMM", "temperature upper 95%", 0.53, tempEffect[["upr"]], 2)

v$gammR2 <- f(100 * attr(r2Top, "adj.r.squared"), 0)
v$gammWeight <- f(100 * dredged$weight[1], 1)
v$gammDelta <- f(dredged$delta[2])
v$dryWet <- f(dryWet[["est"]])
v$dryWetLwr <- f(dryWet[["lwr"]])
v$dryWetUpr <- f(dryWet[["upr"]])
v$tempEst <- f(tempEffect[["est"]])
v$tempLwr <- f(tempEffect[["lwr"]])
v$tempUpr <- f(tempEffect[["upr"]])
```

::: {.slot}
**Paste here:** Results, *Total canopy production*, second paragraph (accepted manuscript l. 297–307).
:::

```{r key-p2, results = "asis"}
valueKey(tribble(
  ~statistic, ~published, ~name,
  "Adjusted R² of best GAMM (%)", "≈ 75", "gammR2",
  "Weight of best GAMM (%)", "> 90", "gammWeight",
  "ΔAICc to next model", "5.47", "gammDelta",
  "Non-swamp/swamp multiplier", "2.01", "dryWet",
  "  lower 95%", "1.57", "dryWetLwr",
  "  upper 95%", "2.57", "dryWetUpr",
  "Temperature coefficient", "0.40", "tempEst",
  "  lower 95%", "0.27", "tempLwr",
  "  upper 95%", "0.53", "tempUpr"))
```

```{r fig2, fig.width = 9, fig.height = 4.5, fig.cap = "**Fig. 2**"}
plotProduction |>
  pivot_longer(c(all, bolian, bolianba), names_to = "measure") |>
  summarise(mean = mean(value), se = seMean(value),
            .by = c(measure, Condition)) |>
  mutate(
    panel = if_else(measure == "bolianba", "(b)", "(a)"),
    measure = recode(measure, all = "All leaves",
                     bolian = "Excluding liana species",
                     bolianba = "Excluding liana species"),
    Condition = recode(Condition, D = "Non-swamp", W = "Swamp")) |>
  ggplot(aes(Condition, mean, fill = measure)) +
  geom_col(position = position_dodge(0.9), width = 0.8) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                position = position_dodge(0.9), width = 0.2) +
  facet_wrap(~ panel) +
  scale_fill_manual(values = c("#40798C", "#9EC1A3")) +
  labs(x = NULL, fill = NULL,
       y = expression(Canopy~production~(g~year^-1~m^-2))) +
  theme_bw() +
  theme(legend.position = "bottom")
```

::: {.slot}
**Paste here:** caption of Fig. 2 (accepted manuscript l. 678–682). Panel (b) is per square metre of basal area (g year⁻¹ m⁻² m⁻²).
:::

```{r fig3, fig.width = 8, fig.height = 9, fig.cap = "**Fig. 3**"}
newDates <- sort(unique(tll$date1))
gammPred <- map(c(W = "W", D = "D"), \(cond) {
  nd <- tibble(date1 = newDates, Condition = cond,
               temp = tll$temp[match(newDates, tll$date1)])
  p <- predict(best$gam, newdata = nd, se.fit = TRUE)
  mutate(nd, fit = p$fit, se = p$se.fit)
}) |>
  list_rbind() |>
  mutate(Condition = recode(Condition, W = "Wet", D = "Dry"))

fig3a <- ggplot(gammPred, aes(date1, fit, colour = Condition,
                              fill = Condition)) +
  geom_ribbon(aes(ymin = fit - 1.96 * se, ymax = fit + 1.96 * se),
              colour = NA, alpha = 0.3) +
  geom_line(aes(linetype = Condition), linewidth = 1.2) +
  geom_point(data = mutate(tll, Condition = recode(Condition, W = "Wet",
                                                   D = "Dry")),
             aes(y = log(prod), shape = Condition), alpha = 0.7) +
  scale_y_continuous(breaks = log(c(4, 8, 16, 32, 64, 128)),
                     labels = c(4, 8, 16, 32, 64, 128)) +
  scale_colour_manual(values = c(Wet = "#1F363D", Dry = "#8D8266")) +
  scale_fill_manual(values = c(Wet = "#40798C", Dry = "#9EC1A3")) +
  labs(x = NULL, title = "(a)", colour = "Plot condition",
       fill = "Plot condition", linetype = "Plot condition",
       shape = "Plot condition",
       y = expression("Basal area-normalized litter production"~
                        (g~fortnight^-1~m^-2~m^-2))) +
  theme_bw()
fig3b <- tll |>
  distinct(date1, temp) |>
  ggplot(aes(date1, temp)) +
  geom_line(linewidth = 1) +
  labs(x = "Date", y = "Mean daily temperature (°C)", title = "(b)") +
  theme_bw()
patchwork::wrap_plots(fig3a, fig3b, ncol = 1, heights = c(2, 1))
```

::: {.slot}
**Paste here:** caption of Fig. 3 (accepted manuscript l. 683–688). See Note 5 on panel (b).
:::

## Species canopy productivity

```{r litter-consol, include = FALSE}
# Identified tree species in the litter (41). "Vatica pauciflora " has a
# trailing space in the data.
sppSel <- c(
  "Adenanthera malayana", "Alstonia pneumatophora", "Aporosa falcifera",
  "Aporosa frutescens", "Aporosa prainiana", "Archidendron clypearia",
  "Artocarpus kemando", "Bhesa paniculata", "Buchanania sessifolia",
  "Calophyllum tetrapterum var. tetrapterum",
  "Calophyllum wallichianum var. incrassatum", "Campnosperma squamatum",
  "Canarium pilosum", "Diospyros maingayi", "Dysoxylum cauliflorum",
  "Elaeocarpus mastersii", "Elaeocarpus stipularis", "Gironniera nervosa",
  "Gluta wallichii", "Gynotroches axillaris", "Hevea brasiliensis",
  "Knema malayana", "Lindera lucida", "Lithocarpus ewyckii",
  "Litsea firma", "Lophopetalum multinervium", "Maasia glauca",
  "Madhuca tomentosa", "Mallotus paniculatus", "Myristica cinnamomea",
  "Nothaphoebe umbelliflora", "Pentace triptera", "Pometia pinnata",
  "Prunus polystachya", "Pternandra echinata", "Rhodamnia cinerea",
  "Sterculia cordata", "Sterculia macrophylla", "Timonius wallichianus",
  "Vatica pauciflora ", "Xanthophyllum flavescens")

# Survey duration per species, over all its collections: from first to
# last collection, or from the only collection to the end of the survey.
lastDate <- max(lamina$date1, na.rm = TRUE)
duration <- lamina |>
  filter(species %in% sppSel) |>
  summarise(start = min(date1, na.rm = TRUE),
            end = max(date1, na.rm = TRUE),
            .by = species) |>
  mutate(duration = as.numeric(if_else(end == start,
                                       lastDate - start, end - start)))

# All plot x species combinations for the selected species that occur as
# trees somewhere in the ten plots (36 species). A litter sum containing
# an NA becomes 0, as in the historical tapply() + is.na() <- 0.
litterSppPlot <- expand_grid(plot = sort(unique(lamina$Plot)),
                             species = intersect(sppSel, tree$species)) |>
  left_join(summarise(lamina, litter.collection = sum(Dry.Mass),
                      .by = c(Plot, species)),
            by = c(plot = "Plot", "species")) |>
  left_join(summarise(tree, ten.ba = sum(ba), .by = c(plot, species)),
            by = c("plot", "species")) |>
  replace_na(list(litter.collection = 0, ten.ba = 0)) |>
  left_join(select(duration, species, start, duration), by = "species") |>
  # Litter traps total 4 m2 per plot: g per m2 of plot area per year
  mutate(litter.production = litter.collection / 4 / (duration / 365),
         code = toupper(paste0(substr(word(species, 1), 1, 1),
                               substr(word(species, 2), 1, 2))))

# Drop species-plots with no litter, then those without the species as a
# tree. HISTORICAL QUIRK: the Q6 Madhuca trees are named "Madhuca sp.", so
# the Q6 "Madhuca tomentosa" row has ten.ba = 0 and is dropped here.
litterSppPlot <- litterSppPlot |>
  filter(litter.production != 0) |>
  filter(ten.ba != 0)
```

```{r siv-cn, include = FALSE}
ssi <- read.csv("../ecophysio_traits/raw_data/SSI Jan21.csv")
# "Madhuca tomentosa" is absent from the SSI file, which has only
# "Madhuca sp."; the historical code gives every unmatched species the
# "Madhuca sp." values.
sivMissing <- setdiff(litterSppPlot$species, ssi$X)
madhucaSsi <- filter(ssi, X == "Madhuca sp.")

# HISTORICAL ERROR: C:N per species is the mean of the per-sample ratios
# over ALL rows of CHNS v3.csv, i.e. fresh AND senesced leaves.
chns <- bind_rows(read.csv("data/CHNS v3_senesced.csv"),
                  read.csv("data/CHNS v3_fresh.csv"))
cnHist <- summarise(chns, CNratio = mean(Ratio), .by = Species)

litterSppPlot <- litterSppPlot |>
  left_join(select(ssi, X, SSI = ssi.ba, iv.wet), by = c(species = "X")) |>
  mutate(SSI = replace_na(SSI, madhucaSsi$ssi.ba),
         iv.wet = replace_na(iv.wet, madhucaSsi$iv.wet)) |>
  left_join(cnHist, by = c(species = "Species"))

# Compare with the historical intermediate output
feb21 <- read.csv("data/litter.production w species-plot CAA Feb21.csv")
feb21Diff <- litterSppPlot |>
  inner_join(feb21, by = c("plot", "species"), suffix = c("", "_feb21"))
feb21MaxDiff <- c("litter.collection", "ten.ba", "duration",
                  "litter.production", "SSI", "iv.wet", "CNratio") |>
  set_names() |>
  map_dbl(\(x) max(abs(feb21Diff[[x]] - feb21Diff[[paste0(x, "_feb21")]]),
                   na.rm = TRUE))
addCheck("Species-plot data", "rows (Feb21 CSV)", nrow(feb21),
         nrow(litterSppPlot), 0)
addCheck("Species-plot data", "rows matched to Feb21 CSV", nrow(feb21),
         nrow(feb21Diff), 0)
addCheck("Species-plot data", "max abs difference, all columns", 0,
         max(feb21MaxDiff), 6)
```

```{r traits, include = FALSE}
# HJ: per-leaf traits; "#DIV/0!" cells become NA. LDMC is in mg g-1, so
# it is divided by 1000 to match PJ's g g-1.
traitsHj <- read.csv("data/HJ leaf data full.csv") |>
  mutate(across(c(thickness, ldmc, SLA),
                \(x) suppressWarnings(as.numeric(x))),
         code = toupper(substr(indiv, 1, 3))) |>
  summarise(SLA = mean(SLA, na.rm = TRUE),
            LDMC = mean(ldmc, na.rm = TRUE) / 1000,
            LT = mean(thickness, na.rm = TRUE),
            .by = code)

# PJ: leaf thickness is the mean of three measurements per leaf; note that
# the species mean of LT does not drop NAs (historical)
traitsPj <- read.csv("data/PJ leaf dat.csv") |>
  mutate(meanlt = rowMeans(across(starts_with("MESOPHYLL_THICKNESS")),
                           na.rm = TRUE)) |>
  summarise(SLA = mean(SLA_cm2.g.1, na.rm = TRUE),
            LDMC = mean(LDMC, na.rm = TRUE),
            LT = mean(meanlt),
            .by = SPECIES) |>
  rename(code = SPECIES)

# Species with traits AND litter data; C:N is the (mixed) species mean
traitsOnly <- bind_rows(traitsHj, traitsPj) |>
  inner_join(summarise(litterSppPlot, CNratio = mean(CNratio),
                       .by = code),
             by = "code") |>
  arrange(code)

leafPca <- prcomp(select(traitsOnly, SLA, LDMC, LT, CNratio),
                  scale. = TRUE)
pcaVariance <- summary(leafPca)$importance["Proportion of Variance", ]

addCheck("PCA", "species in PCA", 25, nrow(traitsOnly), 0)
addCheck("PCA", "PC1 variance (%)", 48.3, 100 * pcaVariance[["PC1"]], 1)
addCheck("PCA", "PC2 variance (%)", 38.7, 100 * pcaVariance[["PC2"]], 1)
```

```{r model-data, include = FALSE}
# Species-plot rows of the 25 species with traits, plus PCA scores. The
# historical chain of merge() calls is kept because the row order it
# leaves matters: the lmer refits in the d-separation tests are sensitive
# to it (Fisher's C of rank 5 changes in the third decimal).
mergeHist <- function(x, trait) {
  rbind(merge(x, as.data.frame(traitsHj[c("code", trait)]), by = "code"),
        merge(x, as.data.frame(traitsPj[c("code", trait)]), by = "code"))
}
litter <- litterSppPlot |>
  arrange(species, plot) |>
  as.data.frame() |>
  mergeHist("SLA") |>
  mergeHist("LT") |>
  mergeHist("LDMC") |>
  merge(data.frame(code = traitsOnly$code, leafPca$x), by = "code") |>
  as_tibble()

# HISTORICAL QUIRK: after the PCA, CTE's C:N is overwritten with the mean
# over species ("assign CTE the mean value for now"), so CTE enters the
# C:N models with this value but the PCA with its own.
cnCteHist <- traitsOnly$CNratio[traitsOnly$code == "CTE"]
litter$CNratio[litter$code == "CTE"] <- mean(traitsOnly$CNratio)

litter <- litter |>
  mutate(
    loglitter = log(litter.production),
    logba = log(ten.ba),
    # Weight each species-plot by 1 / (number of plots of that species)
    weights = 1 / n(), .by = species) |>
  mutate(
    plotwet = as.numeric(plot %in% c("Q3", "Q4", "Q6", "Q9", "Q10")),
    SLA = SLA / 100,
    # Undocumented in the paper: SIV enters every model square-rooted
    iv.wet = sqrt(iv.wet))

addCheck("Path-model data", "species", 25, n_distinct(litter$species), 0)
addCheck("Path-model data", "species-plot rows", 69, nrow(litter), 0)
```

```{r component-models, include = FALSE}
library(lme4)
library(piecewiseSEM)

traitNames <- c("SLA", "LT", "LDMC", "CNratio", "PC1", "PC2")
plotTypes <- c(noPlot = FALSE, plot = TRUE)

# The formula is spliced into the call (bquote) so that piecewiseSEM can
# refit the models for its d-separation tests.
fitLm <- function(fm) {
  eval(bquote(lm(.(fm), weights = weights, data = litter)))
}
fitLmer <- function(fixed, plotType) {
  fm <- reformulate(c("logba", fixed, if (plotType) "plotwet",
                      "(1 | plot)", "(logba | species)"),
                    response = "loglitter")
  eval(bquote(lmer(.(fm), data = litter)))
}

# Trait <-> SIV models (S: SIV -> trait; D: trait -> SIV), and the nulls
modS0 <- map(set_names(traitNames), \(t) fitLm(reformulate("1", t)))
modS <- map(set_names(traitNames), \(t) fitLm(reformulate("iv.wet", t)))
modD <- map(set_names(traitNames), \(t) fitLm(reformulate(t, "iv.wet")))
modD0 <- fitLm(iv.wet ~ 1)

# Canopy production (log litter production ~ log BA + ...), without and
# with plot type
modL0 <- map(plotTypes, \(p) fitLmer(NULL, p))
modL_siv <- map(plotTypes, \(p) fitLmer("iv.wet", p))
modL_trait <- map(plotTypes, \(p)
  map(set_names(traitNames), \(t) fitLmer(t, p)))
modL_sivTrait <- map(plotTypes, \(p)
  map(set_names(traitNames), \(t) fitLmer(c("iv.wet", t), p)))

# Published coefficients and 95% CIs. The published CIs are Wald (normal)
# intervals, estimate +/- 1.96 SE, for the lm models too (t-based
# intervals differ in the second decimal).
coefCi <- function(mod, term) {
  cf <- coef(summary(mod))[term, ]
  cf[["Estimate"]] + c(est = 0, lwr = -1.96, upr = 1.96) * cf[["Std. Error"]]
}
componentCoefs <- tribble(
  ~path,           ~mod,             ~term,     ~pub,
  "SLA -> SIV",    modD$SLA,         "SLA",     c(-0.29, -0.41, -0.17),
  "PC1 -> SIV",    modD$PC1,         "PC1",     c(0.07, 0.04, 0.11),
  "LT -> SIV",     modD$LT,          "LT",      c(1.20, 0.44, 1.95),
  "C:N -> SIV",    modD$CNratio,     "CNratio", c(0.009, 0.004, 0.013),
  "SIV -> canopy", modL_siv$noPlot,  "iv.wet",  c(-1.67, -4.37, 1.04),
  "BA -> canopy",  modL_siv$noPlot,  "logba",   c(0.93, 0.61, 1.25)
) |>
  mutate(repro = map2(mod, term, coefCi))

for (i in seq_len(nrow(componentCoefs))) {
  digits <- if (componentCoefs$term[i] == "CNratio") 3 else 2
  for (j in 1:3) {
    addCheck("Component models",
             paste(componentCoefs$path[i], c("estimate", "lower", "upper")[j]),
             componentCoefs$pub[[i]][j], componentCoefs$repro[[i]][j], digits)
  }
}
componentR2 <- c(SLA = 24.10, PC1 = 21.82, LT = 12.49, CNratio = 18.25,
                 PC2 = 6.06)
for (t in names(componentR2)) {
  addCheck("Component models", paste(t, "-> SIV R² (%)"), componentR2[[t]],
           100 * summary(modD[[t]])$r.squared, 2)
}
addCheck("Component models", "SIV -> LDMC R² (%)", 3.77,
         100 * summary(modS$LDMC)$r.squared, 2)
```

```{r sem-list, include = FALSE}
# One SEM per trait for each hypothesis, in the historical order. Model
# components follow the historical psem() argument order.
buildSems <- function(p) {
  list(
    i    = map(traitNames, \(t) psem(modL_sivTrait[[p]][[t]], modD[[t]])),
    ii   = map(traitNames, \(t) psem(modL_sivTrait[[p]][[t]], modS[[t]])),
    vi   = map(traitNames, \(t) psem(modD[[t]], modL_siv[[p]])),
    vii  = map(traitNames, \(t) psem(modS[[t]], modL_trait[[p]][[t]])),
    iii  = map(traitNames, \(t)
      psem(modL_sivTrait[[p]][[t]], modS0[[t]], modD0)),
    iv   = map(traitNames, \(t) psem(modL_siv[[p]], modD0, modS0[[t]])),
    viii = map(traitNames, \(t) psem(modD[[t]], modS0[[t]], modL0[[p]])),
    ix   = map(traitNames, \(t) psem(modS[[t]], modD0, modL0[[p]])),
    v    = map(traitNames, \(t)
      psem(modL_trait[[p]][[t]], modS0[[t]], modD0)))
}
sems <- map(names(plotTypes), \(p) {
  s <- buildSems(p)
  tibble(plotType = p, BH = rep(names(s), each = length(traitNames)),
         trait = rep(traitNames, length(s)),
         sem = unname(list_flatten(s)))
}) |>
  list_rbind() |>
  mutate(id = row_number(), .before = 1)

# HISTORICAL BUG: model 55 (plot type, BH i, SLA) was written as
# psem(L8.1ct, S1c) instead of psem(L8.1ct, D1c), so it duplicates the
# BH ii model 61 but is labelled BH i.
idBug <- with(sems, id[plotType == "plot" & BH == "i" & trait == "SLA"])
sems$sem[[idBug]] <- psem(modL_sivTrait$plot$SLA, modS$SLA)
```

```{r sem-selection, include = FALSE}
# AICc for every SEM; a model that fails becomes NA and drops silently
# (historical tryCatch)
sems$AICc <- map_dbl(sems$sem, \(m)
  tryCatch(AIC(m, aicc = TRUE), error = \(e) NA_real_))
sems <- sems |>
  mutate(dAICc = AICc - min(AICc, na.rm = TRUE),
         # Fig. 5 weights: from unrounded dAICc, relative likelihoods
         # rounded to 5 dp, weights to 3 dp (historical rounding)
         RL = round(exp(-0.5 * dAICc), 5),
         weight_sow = round(RL / sum(RL), 3)) |>
  arrange(dAICc) |>
  # Table 1 weights: from dAICc rounded to 2 dp (historical rounding)
  mutate(rank = row_number(),
         dAICc_tab = round(dAICc, 2),
         weight = round(round(exp(-0.5 * dAICc_tab), 5) /
                          sum(round(exp(-0.5 * dAICc_tab), 5),
                              na.rm = TRUE), 3))

# R² (marginal, %) per component and Fisher's C. The historical code read
# Fisher's C from summary(), which fails under piecewiseSEM 2.1.2 for
# SEMs with an intercept-only component. For those, dSep() + fisherC() is
# used instead; it can differ from summary() in the third decimal.
sems <- sems |>
  mutate(
    R2 = map_chr(sem, \(m) tryCatch(
      paste(round(100 * rsquared(m)$Marginal, 2), collapse = " / "),
      error = \(e) NA_character_)),
    FisherC_source = "summary",
    FisherC = map_dbl(sem, \(m) tryCatch(
      as.numeric(summary(m, .progressBar = FALSE)$Cstat[1]),
      error = \(e) NA_real_)),
    FisherC_source = if_else(is.na(FisherC), "fisherC", FisherC_source),
    FisherC = if_else(
      is.na(FisherC),
      map_dbl(sem, \(m) tryCatch(
        fisherC(dSep(m, .progressBar = FALSE))$Fisher.C,
        error = \(e) NA_real_)),
      FisherC),
    formula = map_chr(sem, \(m) {
      comps <- m[map_lgl(m, \(x) inherits(x, c("lm", "merMod")))]
      map_chr(comps, \(x) deparse1(formula(x))) |>
        str_remove(fixed(" + (1 | plot) + (logba | species)")) |>
        paste(collapse = "; ")
    }))

sumOfWeights <- sems |>
  summarise(SoW = sum(weight_sow), .by = BH) |>
  arrange(desc(SoW))
```

```{r table1-checks, include = FALSE}
table1Pub <- tribble(
  ~rank, ~BH,   ~R2,             ~FisherC, ~AICc, ~dAICc, ~weight,
  1,     "vi",  "35.55 / 24.10", 0.383,    27.09, 0.00,   0.313,
  2,     "vi",  "35.55 / 21.82", 1.469,    28.41, 1.32,   0.162,
  3,     "i",   "34.86 / 24.10", 0.173,    29.78, 2.69,   0.082,
  4,     "vii", "33.21 / 3.77",  2.849,    30.08, 2.98,   0.071,
  5,     "vi",  "35.55 / 12.49", 3.027,    30.30, 3.20,   0.063,
  6,     "i",   "36.36 / 21.82", 0.627,    30.34, 3.25,   0.062,
  7,     "ii",  "35.37 / 3.77",  0.743,    30.49, 3.39,   0.057,
  8,     "vi",  "35.55 / 18.25", 3.651,    31.05, 3.96,   0.043,
  9,     "vi",  "35.55 / 6.06",  4.758,    32.39, 5.30,   0.022,
  10,    "i",   "35.98 / 12.49", 2.453,    32.59, 5.50,   0.020,
  25,    "iv",  "33.40 / 18.25", NA,       38.61, 11.51,  0.001)

table1Cmp <- table1Pub |>
  left_join(select(sems, rank, BH, R2, FisherC, AICc, dAICc, weight),
            by = "rank", suffix = c("_pub", ""))
for (i in seq_len(nrow(table1Cmp))) {
  r <- table1Cmp[i, ]
  lab <- paste("rank", r$rank)
  addCheck("Table 1", paste(lab, "AICc"), r$AICc_pub, r$AICc, 2)
  addCheck("Table 1", paste(lab, "dAICc"), r$dAICc_pub, r$dAICc, 2)
  addCheck("Table 1", paste(lab, "weight"), r$weight_pub, r$weight, 3)
  if (!is.na(r$FisherC_pub)) {
    addCheck("Table 1", paste(lab, "Fisher's C"), r$FisherC_pub,
             r$FisherC, 3)
  }
}
# R²: Table 1 lists the canopy model first, whatever the component order,
# so the values are compared as sets
table1R2Match <- with(table1Cmp,
  map2_lgl(R2_pub, R2, \(a, b) {
    pa <- sort(as.numeric(str_split_1(a, " / ")))
    pb <- sort(as.numeric(str_split_1(b, " / ")))
    identical(pa, pb)
  }))
bhMatch <- table1Cmp$BH_pub == table1Cmp$BH

sowPub <- c(vi = 61.3, i = 19.2, vii = 9.3, ii = 7.8)
for (h in names(sowPub)) {
  addCheck("Fig. 5", paste("summed weight (%), BH", h), sowPub[[h]],
           100 * sumOfWeights$SoW[sumOfWeights$BH == h], 1)
}
```

```{r values-species, include = FALSE}
ci2 <- function(path, digits = 2) {
  x <- componentCoefs$repro[[which(componentCoefs$path == path)]]
  list(est = f(x[["est"]], digits), lwr = f(x[["lwr"]], digits),
       upr = f(x[["upr"]], digits), raw = x[["est"]])
}
sla <- ci2("SLA -> SIV")
siv <- ci2("SIV -> canopy")
pc1 <- ci2("PC1 -> SIV")
lt <- ci2("LT -> SIV")
cn <- ci2("C:N -> SIV", 3)
ba <- ci2("BA -> canopy")
rankOf <- function(bh, tr, pt = "noPlot") {
  sems$rank[sems$BH == bh & sems$trait == tr & sems$plotType == pt]
}
firstPlotType <- slice_min(filter(sems, plotType == "plot"), rank)

v$slaEst <- sla$est; v$slaLwr <- sla$lwr; v$slaUpr <- sla$upr
v$sivEst <- siv$est; v$sivLwr <- siv$lwr; v$sivUpr <- siv$upr
v$slaIndirect <- f(sla$raw * siv$raw)
v$pc1Est <- pc1$est; v$pc1Lwr <- pc1$lwr; v$pc1Upr <- pc1$upr
v$pc1Indirect <- f(pc1$raw * siv$raw)
v$ltEst <- lt$est; v$ltLwr <- lt$lwr; v$ltUpr <- lt$upr
v$ltIndirect <- f(lt$raw * siv$raw)
v$cnEst <- cn$est; v$cnLwr <- cn$lwr; v$cnUpr <- cn$upr
v$cnIndirect <- f(cn$raw * siv$raw)
v$rankPc1 <- as.character(rankOf("vi", "PC1"))
v$rankLt <- as.character(rankOf("vi", "LT"))
v$rankCn <- as.character(rankOf("vi", "CNratio"))
v$rankPc2 <- as.character(rankOf("vi", "PC2"))
v$sowVi <- f(100 * sumOfWeights$SoW[sumOfWeights$BH == "vi"], 1)
v$baEst <- ba$est; v$baLwr <- ba$lwr; v$baUpr <- ba$upr
v$plotTypeDelta <- f(firstPlotType$dAICc_tab)
v$plotTypeWeight <- f(firstPlotType$weight, 3)
```

::: {.slot}
**Paste here:** Results, *Species canopy productivity*, first paragraph (accepted manuscript l. 310–318).
:::

```{r key-p3, results = "asis"}
valueKey(tribble(
  ~statistic, ~published, ~name,
  "SLA → SIV coefficient", "−0.29", "slaEst",
  "  lower 95%", "−0.41", "slaLwr",
  "  upper 95%", "−0.17", "slaUpr",
  "SIV → canopy coefficient", "−1.67", "sivEst",
  "  lower 95%", "−4.37", "sivLwr",
  "  upper 95%", "1.04", "sivUpr",
  "SLA indirect effect", "0.48", "slaIndirect"))
```

::: {.slot}
**Paste here:** Results, *Species canopy productivity*, second paragraph (accepted manuscript l. 319–329). See Note 4 on the rank given for the C:N model.
:::

```{r key-p4, results = "asis"}
valueKey(tribble(
  ~statistic, ~published, ~name,
  "Rank of PC1 model", "2", "rankPc1",
  "Rank of LT model", "5", "rankLt",
  "Rank of C:N model", "7 (text); 8 (Table 1)", "rankCn",
  "Rank of PC2 model", "9", "rankPc2",
  "PC1 → SIV coefficient", "0.07", "pc1Est",
  "  lower 95%", "0.04", "pc1Lwr",
  "  upper 95%", "0.11", "pc1Upr",
  "PC1 indirect effect", "−0.12", "pc1Indirect",
  "LT → SIV coefficient", "1.20", "ltEst",
  "  lower 95%", "0.44", "ltLwr",
  "  upper 95%", "1.95", "ltUpr",
  "LT indirect effect", "−1.99", "ltIndirect",
  "C:N → SIV coefficient", "0.009", "cnEst",
  "  lower 95%", "0.004", "cnLwr",
  "  upper 95%", "0.013", "cnUpr",
  "C:N indirect effect", "−0.01", "cnIndirect",
  "Summed weight, hypothesis vi (%)", "61.3", "sowVi"))
```

::: {.slot}
**Paste here:** Results, *Species canopy productivity*, third paragraph (accepted manuscript l. 330–337).
:::

```{r key-p5, results = "asis"}
valueKey(tribble(
  ~statistic, ~published, ~name,
  "Basal area coefficient", "0.93", "baEst",
  "  lower 95%", "0.61", "baLwr",
  "  upper 95%", "1.25", "baUpr",
  "ΔAICc of best plot-type model", "11.51", "plotTypeDelta",
  "Weight of best plot-type model", "0.001", "plotTypeWeight"))
```

```{r table1}
# Laid out as in the published Table 1: canopy-production model first,
# variables named as in the paper
prettyTerms <- c("loglitter" = "Canopy production", "logba" = "BA",
                 "iv\\.wet" = "SIV", "plotwet" = "Plot type")
canopyFirst <- function(x, isCanopy) c(x[isCanopy], x[!isCanopy])
sems |>
  filter(rank <= 10 | rank == 25) |>
  mutate(
    parts = str_split(formula, "; "),
    r2parts = str_split(R2, " / "),
    isCanopy = map(parts, \(p) str_starts(p, "loglitter")),
    Models = map2_chr(parts, isCanopy, \(p, c)
      paste(str_replace_all(canopyFirst(p, c), prettyTerms),
            collapse = "<br>")),
    `R² (%)` = map2_chr(r2parts, isCanopy, \(r, c)
      paste(sprintf("%.2f", as.numeric(canopyFirst(r, c))),
            collapse = "<br>"))) |>
  transmute(Rank = rank, Models, `Biological hypothesis` = BH, `R² (%)`,
            `Fisher's C` = sprintf("%.3f", FisherC),
            AICc = sprintf("%.2f", AICc), `ΔAICc` = sprintf("%.2f", dAICc_tab),
            Weight = sprintf("%.3f", weight)) |>
  kable(escape = FALSE, caption = "**Table 1**") |>
  kable_styling(font_size = 12)
```

::: {.slot}
**Paste here:** caption of Table 1 (accepted manuscript l. 658–664). See Note 4 on rank 25 and on Fisher's C.
:::

```{r fig4, fig.width = 7, fig.height = 11, fig.cap = "**Fig. 4**"}
pcaScores <- bind_cols(code = traitsOnly$code, as_tibble(leafPca$x))
pcaArrows <- as_tibble(leafPca$rotation, rownames = "trait") |>
  mutate(trait = recode(trait, CNratio = "C:N"))
fig4a <- ggplot(pcaScores, aes(PC1, PC2)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_text(aes(label = code)) +
  geom_segment(data = pcaArrows,
               aes(x = 0, y = 0, xend = 3 * PC1, yend = 3 * PC2),
               colour = "forestgreen", linewidth = 1,
               arrow = arrow(length = unit(0.2, "cm"))) +
  geom_text(data = pcaArrows, aes(4 * PC1, 3.5 * PC2, label = trait),
            colour = "forestgreen", size = 5) +
  labs(x = sprintf("PC 1 (%.1f%% of variance)", 100 * pcaVariance[["PC1"]]),
       y = sprintf("PC 2 (%.1f%% of variance)", 100 * pcaVariance[["PC2"]]),
       title = "(a)") +
  theme_bw()
fig4b <- litter |>
  distinct(code, SLA, iv.wet) |>
  ggplot(aes(SLA * 100, iv.wet, label = code)) +
  geom_text() +
  labs(x = expression(SLA~(cm^2~g^-1)),
       y = "Habitat specialization (swamp indicator value)",
       title = "(b)") +
  theme_bw()
patchwork::wrap_plots(fig4a, fig4b, ncol = 1)
```

::: {.slot}
**Paste here:** caption of Fig. 4 (accepted manuscript l. 689–699). See Note 2 on the y-axis of panel (b).
:::

```{r fig5, fig.width = 6, fig.height = 4, fig.cap = "**Fig. 5**"}
sumOfWeights |>
  mutate(BH = fct_reorder(BH, SoW)) |>
  ggplot(aes(SoW * 100, BH)) +
  geom_col(fill = "#40798C") +
  geom_text(aes(label = sprintf("%.1f%%", SoW * 100)), hjust = -0.1,
            size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Summed Akaike weight (%)", y = "Biological hypothesis") +
  theme_bw()
```

::: {.slot}
**Paste here:** caption of Fig. 5 (accepted manuscript l. 700 onwards). The published figure shows path diagrams of the four best-supported hypotheses with their summed weights; this bar chart gives the summed weights of all nine.
:::

# Supplementary Material

The Supplementary Material is not part of the accepted manuscript and was not available while this notebook was written. The tables below are regenerated from the data; their captions and any text go in the slots. See Note 6 for the parts that could not be reproduced here.

## Table S1

```{r table-s1}
# Per species, summed over plots; the duration in weeks (rounded) is used
# for the production per plot and the productivity per m2 of BA
tableS1 <- litterSppPlot |>
  summarise(litter.collection = sum(litter.collection),
            ten.ba = sum(ten.ba), No.of.plots = n(),
            .by = c(species, code, duration)) |>
  mutate(duration = round(duration / 7),
         litter.production = round(
           litter.collection / (4 * No.of.plots) / (duration / 52), 1),
         productivity = round(litter.production / (ten.ba / 10000), 1),
         litter.collection = round(litter.collection, 1),
         ten.ba = round(ten.ba, 1)) |>
  arrange(species)

tableS1 |>
  kable(col.names = c("Species", "Code", "Duration (weeks)",
                      "Litter (g)", "BA (cm²)", "Plots",
                      "Production (g m⁻² yr⁻¹)",
                      "Productivity (g m⁻² yr⁻¹ m⁻² BA)")) |>
  kable_styling(full_width = FALSE, font_size = 12) |>
  scroll_box(height = "400px")
```

::: {.slot}
**Paste here:** caption of Table S1, from the Supplementary Material.
:::

## Appendix B: Table S3

```{r table-s3}
library(MuMIn)
# Canopy production models without BA ("N" models), for Appendix B
fitLmerNoBa <- function(fixed) {
  fm <- reformulate(c(if (is.null(fixed)) "1" else fixed,
                      "(1 | plot)", "(logba | species)"),
                    response = "loglitter")
  eval(bquote(lmer(.(fm), data = litter)))
}
traitsS3 <- c(`1` = "SLA", `2` = "LT", `3` = "LDMC", `4` = "CNratio")
modsS3 <- c(
  list(L0 = modL0$noPlot, L0t = modL0$plot),
  imap(traitsS3, \(t, k)
    list(modL_trait$noPlot[[t]], modL_trait$plot[[t]]) |>
      set_names(paste0("L", k), paste0("L", k, "t"))) |>
    unname() |>
    list_flatten(),
  list(L7 = modL_siv$noPlot),
  set_names(modL_sivTrait$noPlot, paste0("L8.", 1:6, "c")),
  list(N0 = fitLmerNoBa(NULL), N0t = fitLmerNoBa("plotwet")),
  imap(traitsS3, \(t, k)
    list(fitLmerNoBa(t), fitLmerNoBa(c("plotwet", t))) |>
      set_names(paste0("N", k), paste0("N", k, "t"))) |>
    unname() |>
    list_flatten(),
  list(N7 = fitLmerNoBa("iv.wet")),
  map(set_names(traitNames, paste0("N8.", 1:6, "c")), \(t)
    fitLmerNoBa(c("iv.wet", t))))

modelSelS3 <- model.sel(modsS3)
tableS3 <- tibble(model = rownames(modelSelS3)) |>
  bind_cols(
    map(rownames(modelSelS3), \(m) {
      cm <- coef(summary(modsS3[[m]]))
      terms <- c("logba", "plotwet", "SLA", "LT", "LDMC", "CNratio",
                 "iv.wet", "PC1", "PC2")
      map_chr(set_names(terms), \(j)
        if (j %in% rownames(cm)) {
          sprintf("%.2f (±%.2f)", cm[j, "Estimate"], cm[j, "Std. Error"])
        } else NA_character_) |>
        as_tibble_row()
    }) |>
      list_rbind(),
    as.data.frame(modelSelS3)[c("df", "logLik", "AICc", "delta",
                                "weight")] |>
      mutate(across(everything(), as.numeric)) |>
      as_tibble()) |>
  mutate(R2adj = map_dbl(model, \(m) tryCatch(
    attr(r.squaredLR(modsS3[[m]]), "adj.r.squared"),
    error = \(e) NA_real_)))

tableS3 |>
  rename(BA = logba, `Plot type` = plotwet, `C:N` = CNratio,
         SIV = iv.wet, `adj. R²` = R2adj) |>
  kable(digits = 3) |>
  kable_styling(font_size = 11) |>
  scroll_box(height = "400px")
```

::: {.slot}
**Paste here:** Appendix B text and the caption of Table S3, from the Supplementary Material.
:::

# Notes

These notes are not part of the paper.

## Note 1. Inputs {#note-1}

```{r inputs}
tibble(
  File = c("data/Cleaned lamina+PJ.csv", "data/ten plot trees.csv",
           "../ecophysio_traits/raw_data/SSI Jan21.csv",
           "data/CHNS v3_senesced.csv + data/CHNS v3_fresh.csv",
           "data/HJ leaf data full.csv", "data/PJ leaf dat.csv",
           "../../TFERP/LTFEM_aux/data/nea/weather.gov.sg/",
           "data/litter.production w species-plot CAA Feb21.csv",
           "data/data for GAMM.csv"),
  Role = c("Leaf litter per collection, plot and species",
           "Tree DBH in the ten plots (Figshare file 29128734)",
           "Swamp indicator value (SIV) and SSI per species (sister repo)",
           "C, N and C:N per leaf sample; together = historical CHNS v3.csv",
           "Leaf traits, 15 species (HJ)",
           "Leaf traits, 10 species (PJ)",
           "NEA daily weather: S122 Khatib, S69 Upper Peirce, S40 Mandai",
           "Historical intermediate output; used here only as a check",
           "Historical GAMM input; used here only as a check")
) |>
  kable() |>
  kable_styling(full_width = FALSE)
```

The rebuilt species–plot table (`r nrow(litterSppPlot)` rows, `r n_distinct(litterSppPlot$species)` species) matches the historical intermediate file `litter.production w species-plot CAA Feb21.csv` row for row (largest absolute difference `r format(signif(max(feb21MaxDiff), 2), scientific = TRUE)`). The rebuilt GAMM data (`r nrow(gammData)` plot–collections) match `data for GAMM.csv` (largest difference `r format(signif(max(gammMaxDiff), 2), scientific = TRUE)`), with the weather variables recomputed from freshly downloaded NEA records.

## Note 2. Historical choices kept in this reproduction {#note-2}

- **Leaf C:N mixed fresh and senesced leaves.** The Methods state that C and N were measured on fresh leaves, but each species' C:N is the mean over all its CHNS samples, senesced leaves included. This is corrected in `corrected_reanalysis.Rmd`.
- **The *Calophyllum tetrapterum* (CTE) override.** After the PCA, CTE's C:N is replaced by the mean over species, so CTE enters the PCA with C:N = `r f(cnCteHist)` but every C:N model with `r f(mean(traitsOnly$CNratio))`.
- **The duplicated SEM.** Model 55 (plot type, hypothesis i, SLA) pairs the canopy model with SIV → SLA instead of SLA → SIV, duplicating a hypothesis ii model.
- ***Madhuca* naming.** The trees in plot Q6 are named "Madhuca sp." while the litter is "Madhuca tomentosa", so the Q6 *Madhuca* row is lost. `r if (identical(sivMissing, "Madhuca tomentosa")) "*Madhuca tomentosa* is also absent from the SSI file and takes its \"Madhuca sp.\" values." else ""`
- **Square-rooted SIV.** SIV enters every model, and Fig. 4b, square-rooted; the Methods do not say so.
- **Truncated weather windows.** The weather files read stop on `r format(max(weather$date1), "%d %B %Y")`, but collections run to July 2019, so `r nTruncated` collections have truncated 30-day windows. Complete windows leave the GAMM materially unchanged (`audit/09_gamm_fullwindows.R`).

## Note 3. Comparison with the published values {#note-3}

```{r summary-checks}
checks |>
  summarise(checks = n(), matched = sum(match), .by = section) |>
  kable(caption = "Checks by section.") |>
  kable_styling(full_width = FALSE)
```

Of `r nrow(checks)` comparisons with published or historical values, `r sum(checks$match)` match at the published precision. The others:

```{r mismatches}
checks |>
  filter(!match) |>
  kable() |>
  kable_styling(full_width = FALSE)
```

The GAMM effect sizes differ from the published ones in the second decimal place; no confidence-interval method tried (Wald, *t*, profile) recovers them exactly, and package-version drift since 2021 is the likely cause (unverified). The published upper confidence limit for PC1 → SIV, 0.11, is 0.10 in the reproduction, probably a transcription slip.

```{r all-checks}
checks |>
  kable(caption = "All checks.") |>
  kable_styling(full_width = FALSE, font_size = 11) |>
  scroll_box(height = "400px")
```

## Note 4. Table 1 and the text {#note-4}

- The R² values match Table 1 for `r sum(table1R2Match)` of `r length(table1R2Match)` rows, and the hypothesis labels for `r sum(bhMatch)` of `r length(bhMatch)`. At rank 25, the structure (canopy production ~ BA + C:N + plot type; C:N ~ SIV) is hypothesis `r table1Cmp$BH[table1Cmp$rank == 25]`, but Table 1 labels it "iv"; its canopy-model R² is `r str_extract(table1Cmp$R2[table1Cmp$rank == 25], "[0-9.]+$")` against the published 33.40.
- The text calls the C:N model "seventh-ranked", but it is ranked `r v$rankCn` in Table 1; the Discussion calls the LDMC model "sixth-ranked", but it is ranked `r rankOf("ii", "LDMC")`.
- Fisher's C is read from `summary()`, as in the original code. Under piecewiseSEM 2.1.2, `summary()` fails for SEMs with an intercept-only component (`r sum(sems$FisherC_source == "fisherC")` of `r nrow(sems)`); for those, `fisherC(dSep())` is used.
- The path coefficients' 95% confidence intervals are Wald intervals (estimate ± 1.96 SE), which reproduce the published ones.

## Note 5. Fig. 3 {#note-5}

The published caption describes panel (b) as the number of rain days in a month, but the panel shows mean daily temperature over the 30 days before each collection, as here.

## Note 6. Supplementary Material {#note-6}

Appendix A (habitat associations from the 40 vegetation plots) is not reproduced: the swamp indicator values were computed from data that are not in this repository, and are read here from `SSI Jan21.csv`. Supplementary figures were not reproduced.

## Note 7. Reference outputs and software {#note-7}

```{r reference-outputs}
# Written for comparison with corrected_reanalysis.Rmd
dir.create("reference", showWarnings = FALSE)
write_csv(select(litterSppPlot, -start), "reference/historical_species_plot.csv")
write_csv(bind_cols(traitsOnly, as_tibble(leafPca$x)),
          "reference/historical_species_traits_pca.csv")
write_csv(as_tibble(leafPca$rotation, rownames = "trait"),
          "reference/historical_pca_loadings.csv")
write_csv(
  bind_rows(
    imap(modD, \(m, t) tibble(model = paste(t, "-> SIV"),
                              term = names(coef(m)),
                              estimate = unname(coef(m)),
                              r2 = summary(m)$r.squared)) |>
      list_rbind(),
    imap(modS, \(m, t) tibble(model = paste("SIV ->", t),
                              term = names(coef(m)),
                              estimate = unname(coef(m)),
                              r2 = summary(m)$r.squared)) |>
      list_rbind(),
    tibble(model = "SIV -> canopy (L7)",
           term = names(fixef(modL_siv$noPlot)),
           estimate = unname(fixef(modL_siv$noPlot)), r2 = NA)),
  "reference/historical_component_models.csv")
write_csv(select(sems, -sem), "reference/historical_sem_table.csv")
write_csv(sumOfWeights, "reference/historical_sum_of_weights.csv")
write_csv(tableS1, "reference/historical_tableS1.csv")
write_csv(tableS3, "reference/historical_tableS3.csv")
write_csv(
  tibble(quantity = c("weight_top", "dAICc_2nd", "r2adj_top",
                      "dry_wet", "dry_wet_lwr", "dry_wet_upr",
                      "temp", "temp_lwr", "temp_upr"),
         value = c(dredged$weight[1], dredged$delta[2],
                   attr(r2Top, "adj.r.squared"), dryWet, tempEffect)),
  "reference/historical_gamm.csv")
write_csv(checks, "reference/historical_checks.csv")
```

The main outputs are written to `reference/` (prefix `historical_`) as the baseline for the corrected analysis. Analyses were run with piecewiseSEM 2.1.2 (the version current when the paper was written) and gamm4 0.2-6; the paper gives R ver. 3.4.0, which cannot run piecewiseSEM 2.1.2.

```{r session-info}
sessionInfo()
```
