---
title: "Historical reproduction of Lam et al. (2022)"
subtitle: "Habitat adaptation mediates the influence of leaf traits on canopy productivity: evidence from a tropical freshwater swamp forest"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Version-pinned packages (piecewiseSEM 2.1.2, gamm4 0.2-6) live in the
# project-local library .Rlib/ (gitignored); see the README.
.libPaths(c(normalizePath(".Rlib"), .libPaths()))
stopifnot(packageVersion("piecewiseSEM") == "2.1.2")

library(tidyverse)
library(knitr)
library(kableExtra)
```

# Purpose

This notebook reproduces the published results of Lam et al. (2022, *Ecosystems* 25: 1006–1019, https://doi.org/10.1007/s10021-021-00697-7) from the historical input data, as the analysis was actually run. It is a clean rewrite of the historical scripts (commit `74befc2`, kept verbatim in `scripts/historical/`), not a correction of them: every historical choice is kept, including the ones now known to be errors, and each one is flagged where it happens. The corrected reanalysis is in `corrected_reanalysis.Rmd`.

Each published number is checked against the reproduction, and the checks are collected in the [summary](#summary-of-checks) at the end. The main outputs are written to `reference/` as the baseline for the corrected reanalysis.

Historical quirks kept here on purpose:

- **C:N ratio from fresh and senesced leaves mixed.** The published Methods say C and N were measured on fresh leaves, but the species means average every row of the CHNS file, senesced leaves included.
- **The *Calophyllum tetrapterum* (CTE) override.** CTE's C:N is replaced by the mean of all species after the PCA, so CTE carries one value in the PCA and another in every C:N model.
- **The duplicated SEM.** Model 55 (plot-type block, hypothesis i, SLA) pairs the canopy model with the SIV → SLA model instead of SLA → SIV, duplicating a hypothesis ii model.
- ***Madhuca* naming.** The trees in plot Q6 are named "Madhuca sp." while the litter is "Madhuca tomentosa", so the Q6 *Madhuca* row is lost.
- **Truncated weather windows in the GAMM.** The weather files read stop at June 2019, but collections run to July 2019.

# Inputs

```{r inputs}
tibble(
  File = c("data/Cleaned lamina+PJ.csv", "data/ten plot trees.csv",
           "../ecophysio_traits/raw_data/SSI Jan21.csv",
           "data/CHNS v3_senesced.csv + data/CHNS v3_fresh.csv",
           "data/HJ leaf data full.csv", "data/PJ leaf dat.csv",
           "../../TFERP/LTFEM_aux/data/nea/",
           "data/litter.production w species-plot CAA Feb21.csv",
           "data/data for GAMM.csv"),
  Role = c("Leaf litter per collection, plot and species",
           "Tree DBH in the ten plots (Figshare file 29128734)",
           "Swamp indicator value (SIV) and SSI per species (sister repo)",
           "C, N and C:N per leaf sample; together = historical CHNS v3.csv",
           "Leaf traits, 15 species (HJ)",
           "Leaf traits, 10 species (PJ)",
           "NEA daily weather: S122 Khatib, S69 Upper Peirce, S40 Mandai",
           "Historical intermediate output; used here only as a check",
           "Historical GAMM input; used here only as a check")
) |>
  kable() |>
  kable_styling(full_width = FALSE)
```

```{r checks-helper}
# Every comparison with a published (or historical) value is added to
# `checks`, which is summarised at the end. A check passes when the
# reproduced value, rounded to the published precision, equals it.
checks <- tibble()
addCheck <- function(section, item, published, reproduced, digits) {
  checks <<- bind_rows(checks, tibble(
    section = section, item = item, published = published,
    reproduced = round(reproduced, digits),
    match = round(reproduced, digits) == published))
}
```

# Species–plot litter production

This section follows part I of the historical code: litter mass per species and plot, basal area (BA) per species and plot, survey duration, annual litter production, SIV and C:N.

```{r litter-data}
lamina <- read.csv("data/Cleaned lamina+PJ.csv") |>
  mutate(date1 = as.Date(General.date, format = "%d/%m/%Y"),
         date2 = format(date1, "%Y-%m"))

# Rows with NA in any column are dropped (historical na.omit()), and so are
# trees without a species name.
tree <- read.csv("data/ten plot trees.csv") |>
  na.omit() |>
  filter(species != "") |>
  mutate(ba = pi * (dbh_2018 / 2)^2)

# The historical code picked these by factor-level index; the names below
# are what those indices resolve to (verified under R 4.3.3).
lianas <- c(
  "Agelaea borneensis", "Another climber", "Bauhinia semibifida",
  "Dalbergia pseudo-sissoo", "Derris amoena",
  "Fissistigma latifolium var. ovoideum", "Grewia laevigata",
  "Iodes ovalis", "Iodes velutina", "Leuconotis griffithii",
  "Smilax setosa", "Strychnos", "Uncaria cordata",
  "Uncaria cordata var. cordata f. sundaica", "Willughbeia coriacea")
# Identified tree species in the litter (41). "Vatica pauciflora " has a
# trailing space in the data.
sppSel <- c(
  "Adenanthera malayana", "Alstonia pneumatophora", "Aporosa falcifera",
  "Aporosa frutescens", "Aporosa prainiana", "Archidendron clypearia",
  "Artocarpus kemando", "Bhesa paniculata", "Buchanania sessifolia",
  "Calophyllum tetrapterum var. tetrapterum",
  "Calophyllum wallichianum var. incrassatum", "Campnosperma squamatum",
  "Canarium pilosum", "Diospyros maingayi", "Dysoxylum cauliflorum",
  "Elaeocarpus mastersii", "Elaeocarpus stipularis", "Gironniera nervosa",
  "Gluta wallichii", "Gynotroches axillaris", "Hevea brasiliensis",
  "Knema malayana", "Lindera lucida", "Lithocarpus ewyckii",
  "Litsea firma", "Lophopetalum multinervium", "Maasia glauca",
  "Madhuca tomentosa", "Mallotus paniculatus", "Myristica cinnamomea",
  "Nothaphoebe umbelliflora", "Pentace triptera", "Pometia pinnata",
  "Prunus polystachya", "Pternandra echinata", "Rhodamnia cinerea",
  "Sterculia cordata", "Sterculia macrophylla", "Timonius wallichianus",
  "Vatica pauciflora ", "Xanthophyllum flavescens")
```

```{r litter-consol}
# Survey duration per species, over all its collections: from first to
# last collection, or from the only collection to the end of the survey.
lastDate <- max(lamina$date1, na.rm = TRUE)
duration <- lamina |>
  filter(species %in% sppSel) |>
  summarise(start = min(date1, na.rm = TRUE),
            end = max(date1, na.rm = TRUE),
            .by = species) |>
  mutate(duration = as.numeric(if_else(end == start,
                                       lastDate - start, end - start)))

# All plot x species combinations for the selected species that occur as
# trees somewhere in the ten plots (36 species). A litter sum containing
# an NA becomes 0, as in the historical tapply() + is.na() <- 0.
litterSppPlot <- expand_grid(plot = sort(unique(lamina$Plot)),
                             species = intersect(sppSel, tree$species)) |>
  left_join(summarise(lamina, litter.collection = sum(Dry.Mass),
                      .by = c(Plot, species)),
            by = c(plot = "Plot", "species")) |>
  left_join(summarise(tree, ten.ba = sum(ba), .by = c(plot, species)),
            by = c("plot", "species")) |>
  replace_na(list(litter.collection = 0, ten.ba = 0)) |>
  left_join(select(duration, species, start, duration), by = "species") |>
  # Litter traps total 4 m2 per plot: g per m2 of plot area per year
  mutate(litter.production = litter.collection / 4 / (duration / 365),
         code = toupper(paste0(substr(word(species, 1), 1, 1),
                               substr(word(species, 2), 1, 2))))

# Drop species-plots with no litter, then those without the species as a
# tree. HISTORICAL QUIRK: the Q6 Madhuca trees are named "Madhuca sp.", so
# the Q6 "Madhuca tomentosa" row has ten.ba = 0 and is dropped here.
litterSppPlot <- litterSppPlot |>
  filter(litter.production != 0) |>
  filter(ten.ba != 0)
```

```{r siv-cn}
ssi <- read.csv("../ecophysio_traits/raw_data/SSI Jan21.csv")
# "Madhuca tomentosa" is absent from the SSI file, which has only
# "Madhuca sp."; the historical code gives every unmatched species the
# "Madhuca sp." values.
sivMissing <- setdiff(litterSppPlot$species, ssi$X)
madhucaSsi <- filter(ssi, X == "Madhuca sp.")

# HISTORICAL ERROR: C:N per species is the mean of the per-sample ratios
# over ALL rows of CHNS v3.csv, i.e. fresh AND senesced leaves.
chns <- bind_rows(read.csv("data/CHNS v3_senesced.csv"),
                  read.csv("data/CHNS v3_fresh.csv"))
cnHist <- summarise(chns, CNratio = mean(Ratio), .by = Species)

litterSppPlot <- litterSppPlot |>
  left_join(select(ssi, X, SSI = ssi.ba, iv.wet), by = c(species = "X")) |>
  mutate(SSI = replace_na(SSI, madhucaSsi$ssi.ba),
         iv.wet = replace_na(iv.wet, madhucaSsi$iv.wet)) |>
  left_join(cnHist, by = c(species = "Species"))
```

```{r check-feb21}
# Compare with the historical intermediate output
feb21 <- read.csv("data/litter.production w species-plot CAA Feb21.csv")
feb21Diff <- litterSppPlot |>
  inner_join(feb21, by = c("plot", "species"), suffix = c("", "_feb21"))
feb21MaxDiff <- c("litter.collection", "ten.ba", "duration",
                  "litter.production", "SSI", "iv.wet", "CNratio") |>
  set_names() |>
  map_dbl(\(v) max(abs(feb21Diff[[v]] - feb21Diff[[paste0(v, "_feb21")]]),
                   na.rm = TRUE))

addCheck("Species-plot data", "rows (Feb21 CSV)", nrow(feb21),
         nrow(litterSppPlot), 0)
addCheck("Species-plot data", "rows matched to Feb21 CSV", nrow(feb21),
         nrow(feb21Diff), 0)
addCheck("Species-plot data", "max abs difference, all columns", 0,
         max(feb21MaxDiff), 6)
```

The rebuilt table has `r nrow(litterSppPlot)` species–plot rows for `r n_distinct(litterSppPlot$species)` species. It matches the historical intermediate file (`litter.production w species-plot CAA Feb21.csv`) row for row, with a largest absolute difference of `r format(signif(max(feb21MaxDiff), 2), scientific = TRUE)` across litter mass, BA, duration, production, SSI, SIV and C:N. `r if (identical(sivMissing, "Madhuca tomentosa")) "Only *Madhuca tomentosa* is missing from the SSI file, so it is the only species that takes the \"Madhuca sp.\" values." else paste("Species missing from the SSI file:", paste(sivMissing, collapse = ", "))`

# Total canopy production (Fig. 2)

```{r total-production}
# Plot-level stand data for leaf litter of non-liana species (also the
# basis of the GAMM): litter per collection, plot BA, BA-normalised
# production.
plotBa <- summarise(tree, ba = sum(ba), .by = plot)
# aggregate() dropped rows with NA mass or date (historical formula method)
standLitter <- lamina |>
  filter(!species %in% lianas, !is.na(Dry.Mass), !is.na(date1)) |>
  summarise(Dry.Mass = sum(Dry.Mass), .by = c(date1, date2, Plot)) |>
  left_join(distinct(lamina, Plot, Condition), by = "Plot") |>
  left_join(plotBa, by = c(Plot = "plot")) |>
  mutate(prod = Dry.Mass / ba * 100^2 / 4)

# Per plot, in g m-2 yr-1 (the survey lasted about a year): all leaves,
# non-liana leaves, and BA-normalised non-liana leaves
plotProduction <- lamina |>
  summarise(all = sum(Dry.Mass) / 4, .by = Plot) |>
  left_join(summarise(standLitter,
                      bolian = sum(Dry.Mass) / 4,
                      bolianba = sum(prod),
                      .by = c(Plot, Condition)),
            by = "Plot") |>
  arrange(Plot)

seMean <- function(x) sd(x) / sqrt(length(x))
tTests <- c(all = "all", bolian = "bolian", bolianba = "bolianba") |>
  map(\(v) t.test(reformulate("Condition", v), data = plotProduction,
                  var.equal = TRUE))

addCheck("Fig. 2", "mean canopy production (g m-2 yr-1)", 768,
         mean(plotProduction$all), 0)
addCheck("Fig. 2", "SE of canopy production", 48,
         seMean(plotProduction$all), 0)
addCheck("Fig. 2", "total litter (kg)", 30.7,
         sum(lamina$Dry.Mass, na.rm = TRUE) / 1000, 1)
addCheck("Fig. 2", "t, all leaves", 1.33,
         abs(tTests$all$statistic), 2)
addCheck("Fig. 2", "P, all leaves", 0.221, tTests$all$p.value, 3)
addCheck("Fig. 2", "t, excluding lianas", 0.49,
         abs(tTests$bolian$statistic), 2)
addCheck("Fig. 2", "P, excluding lianas", 0.639, tTests$bolian$p.value, 3)
addCheck("Fig. 2", "t, BA-normalised", 6.81,
         abs(tTests$bolianba$statistic), 2)
```

Mean canopy production was `r round(mean(plotProduction$all))` ± `r round(seMean(plotProduction$all))` g m^−2^ yr^−1^ (mean ± SE across the ten plots). Swamp and non-swamp plots did not differ in production of all leaves (*t* = `r round(abs(tTests$all$statistic), 2)`, *P* = `r round(tTests$all$p.value, 3)`) or of non-liana leaves (*t* = `r round(abs(tTests$bolian$statistic), 2)`, *P* = `r round(tTests$bolian$p.value, 3)`), but BA-normalised production was higher in non-swamp plots (*t* = `r round(abs(tTests$bolianba$statistic), 2)`, *P* < 0.001).

```{r fig2, fig.width = 9, fig.height = 4.5, fig.cap = "Figure 2 reproduced. Means ± SE across plots (n = 5 per plot type)."}
plotProduction |>
  pivot_longer(c(all, bolian, bolianba), names_to = "measure") |>
  summarise(mean = mean(value), se = seMean(value),
            .by = c(measure, Condition)) |>
  mutate(
    panel = if_else(measure == "bolianba",
                    "b) BA-normalised, excluding lianas",
                    "a) Canopy production"),
    measure = recode(measure, all = "All leaves",
                     bolian = "Excluding liana species",
                     bolianba = "Excluding liana species"),
    Condition = recode(Condition, D = "Non-swamp", W = "Swamp")) |>
  ggplot(aes(Condition, mean, fill = measure)) +
  geom_col(position = position_dodge(0.9), width = 0.8) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se),
                position = position_dodge(0.9), width = 0.2) +
  facet_wrap(~ panel) +
  scale_fill_manual(values = c("#70A9A1", "#9EC1A3")) +
  labs(x = NULL, fill = NULL,
       y = expression(Production~(g~yr^-1~m^-2)~or~(g~yr^-1~m^-2~m^-2))) +
  theme_bw() +
  theme(legend.position = "bottom")
```

# Table S1

```{r table-s1}
# Per species, summed over plots; the duration in weeks (rounded) is used
# for the production per plot and the productivity per m2 of BA
tableS1 <- litterSppPlot |>
  summarise(litter.collection = sum(litter.collection),
            ten.ba = sum(ten.ba), No.of.plots = n(),
            .by = c(species, code, duration)) |>
  mutate(duration = round(duration / 7),
         litter.production = round(
           litter.collection / (4 * No.of.plots) / (duration / 52), 1),
         productivity = round(litter.production / (ten.ba / 10000), 1),
         litter.collection = round(litter.collection, 1),
         ten.ba = round(ten.ba, 1)) |>
  arrange(species)

tableS1 |>
  kable(col.names = c("Species", "Code", "Duration (weeks)",
                      "Litter (g)", "BA (cm²)", "Plots",
                      "Production (g m⁻² yr⁻¹)",
                      "Productivity (g m⁻² yr⁻¹ m⁻² BA)")) |>
  kable_styling(full_width = FALSE, font_size = 12) |>
  scroll_box(height = "400px")
```

The published Table S1 is in the Supplementary Information, which was not available for checking, so this table has no reference values.

# Leaf traits and PCA (Fig. 4a)

```{r traits}
# HJ: per-leaf traits; "#DIV/0!" cells become NA. LDMC is in mg g-1, so
# it is divided by 1000 to match PJ's g g-1.
traitsHj <- read.csv("data/HJ leaf data full.csv") |>
  mutate(across(c(thickness, ldmc, SLA),
                \(x) suppressWarnings(as.numeric(x))),
         code = toupper(substr(indiv, 1, 3))) |>
  summarise(SLA = mean(SLA, na.rm = TRUE),
            LDMC = mean(ldmc, na.rm = TRUE) / 1000,
            LT = mean(thickness, na.rm = TRUE),
            .by = code)

# PJ: leaf thickness is the mean of three measurements per leaf; note that
# the species mean of LT does not drop NAs (historical)
traitsPj <- read.csv("data/PJ leaf dat.csv") |>
  mutate(meanlt = rowMeans(across(starts_with("MESOPHYLL_THICKNESS")),
                           na.rm = TRUE)) |>
  summarise(SLA = mean(SLA_cm2.g.1, na.rm = TRUE),
            LDMC = mean(LDMC, na.rm = TRUE),
            LT = mean(meanlt),
            .by = SPECIES) |>
  rename(code = SPECIES)

# Species with traits AND litter data; C:N is the (mixed) species mean
traitsOnly <- bind_rows(traitsHj, traitsPj) |>
  inner_join(summarise(litterSppPlot, CNratio = mean(CNratio),
                       .by = code),
             by = "code") |>
  arrange(code)

leafPca <- prcomp(select(traitsOnly, SLA, LDMC, LT, CNratio),
                  scale. = TRUE)
pcaVariance <- summary(leafPca)$importance["Proportion of Variance", ]

addCheck("PCA", "species in PCA", 25, nrow(traitsOnly), 0)
addCheck("PCA", "PC1 variance (%)", 48.3, 100 * pcaVariance[["PC1"]], 1)
addCheck("PCA", "PC2 variance (%)", 38.7, 100 * pcaVariance[["PC2"]], 1)
```

The PCA of four leaf traits (SLA, LDMC, LT and C:N) across `r nrow(traitsOnly)` species explained `r round(100 * pcaVariance[["PC1"]], 1)`% of the variance on PC1 and `r round(100 * pcaVariance[["PC2"]], 1)`% on PC2.

```{r pca-loadings}
leafPca$rotation[, 1:2] |>
  as_tibble(rownames = "trait") |>
  kable(digits = 3, caption = "PCA loadings.") |>
  kable_styling(full_width = FALSE)
```

# Data for the path models

```{r model-data}
# Species-plot rows of the 25 species with traits, plus PCA scores. The
# historical chain of merge() calls is kept because the row order it
# leaves matters: the lmer refits in the d-separation tests are sensitive
# to it (Fisher's C of rank 5 changes in the third decimal).
mergeHist <- function(x, trait) {
  rbind(merge(x, as.data.frame(traitsHj[c("code", trait)]), by = "code"),
        merge(x, as.data.frame(traitsPj[c("code", trait)]), by = "code"))
}
litter <- litterSppPlot |>
  arrange(species, plot) |>
  as.data.frame() |>
  mergeHist("SLA") |>
  mergeHist("LT") |>
  mergeHist("LDMC") |>
  merge(data.frame(code = traitsOnly$code, leafPca$x), by = "code") |>
  as_tibble()

# HISTORICAL QUIRK: after the PCA, CTE's C:N is overwritten with the mean
# over species ("assign CTE the mean value for now"), so CTE enters the
# C:N models with this value but the PCA with its own.
cnCteHist <- traitsOnly$CNratio[traitsOnly$code == "CTE"]
litter$CNratio[litter$code == "CTE"] <- mean(traitsOnly$CNratio)

litter <- litter |>
  mutate(
    loglitter = log(litter.production),
    logba = log(ten.ba),
    # Weight each species-plot by 1 / (number of plots of that species)
    weights = 1 / n(), .by = species) |>
  mutate(
    plotwet = as.numeric(plot %in% c("Q3", "Q4", "Q6", "Q9", "Q10")),
    SLA = SLA / 100,
    # Undocumented in the paper: SIV enters every model square-rooted
    iv.wet = sqrt(iv.wet))

addCheck("Path-model data", "species", 25, n_distinct(litter$species), 0)
addCheck("Path-model data", "species-plot rows", 69, nrow(litter), 0)
```

The path models use `r nrow(litter)` species–plot rows from `r n_distinct(litter$species)` species. CTE enters the PCA with C:N = `r round(cnCteHist, 2)` but every C:N model with `r round(mean(traitsOnly$CNratio), 2)`. SLA is divided by 100, SIV is square-rooted (not stated in the paper), and each row is weighted by the inverse of its species' number of plots.

# Component models

```{r component-models}
library(lme4)
library(piecewiseSEM)

traitNames <- c("SLA", "LT", "LDMC", "CNratio", "PC1", "PC2")
plotTypes <- c(noPlot = FALSE, plot = TRUE)

# The formula is spliced into the call (bquote) so that piecewiseSEM can
# refit the models for its d-separation tests.
fitLm <- function(f) {
  eval(bquote(lm(.(f), weights = weights, data = litter)))
}
fitLmer <- function(fixed, plotType) {
  f <- reformulate(c("logba", fixed, if (plotType) "plotwet",
                     "(1 | plot)", "(logba | species)"),
                   response = "loglitter")
  eval(bquote(lmer(.(f), data = litter)))
}

# Trait <-> SIV models (S: SIV -> trait; D: trait -> SIV), and the nulls
modS0 <- map(set_names(traitNames), \(t) fitLm(reformulate("1", t)))
modS <- map(set_names(traitNames), \(t) fitLm(reformulate("iv.wet", t)))
modD <- map(set_names(traitNames), \(t) fitLm(reformulate(t, "iv.wet")))
modD0 <- fitLm(iv.wet ~ 1)

# Canopy production (log litter production ~ log BA + ...), without and
# with plot type
modL0 <- map(plotTypes, \(p) fitLmer(NULL, p))
modL_siv <- map(plotTypes, \(p) fitLmer("iv.wet", p))
modL_trait <- map(plotTypes, \(p)
  map(set_names(traitNames), \(t) fitLmer(t, p)))
modL_sivTrait <- map(plotTypes, \(p)
  map(set_names(traitNames), \(t) fitLmer(c("iv.wet", t), p)))
```

```{r component-checks}
# Published coefficients and 95% CIs. The published CIs are Wald (normal)
# intervals, estimate +/- 1.96 SE, for the lm models too (t-based
# intervals differ in the second decimal).
coefCi <- function(mod, term) {
  cf <- coef(summary(mod))[term, ]
  cf[["Estimate"]] + c(est = 0, lwr = -1.96, upr = 1.96) * cf[["Std. Error"]]
}
componentCoefs <- tribble(
  ~path,         ~mod,                        ~term,     ~pub,
  "SLA -> SIV",  modD$SLA,                    "SLA",     c(-0.29, -0.41, -0.17),
  "PC1 -> SIV",  modD$PC1,                    "PC1",     c(0.07, 0.04, 0.11),
  "LT -> SIV",   modD$LT,                     "LT",      c(1.20, 0.44, 1.95),
  "C:N -> SIV",  modD$CNratio,                "CNratio", c(0.009, 0.004, 0.013),
  "SIV -> canopy", modL_siv$noPlot,           "iv.wet",  c(-1.67, -4.37, 1.04),
  "BA -> canopy",  modL_siv$noPlot,           "logba",   c(0.93, 0.61, 1.25)
) |>
  mutate(repro = map2(mod, term, coefCi))

for (i in seq_len(nrow(componentCoefs))) {
  digits <- if (componentCoefs$term[i] == "CNratio") 3 else 2
  for (j in 1:3) {
    addCheck("Component models",
             paste(componentCoefs$path[i], c("estimate", "lower", "upper")[j]),
             componentCoefs$pub[[i]][j], componentCoefs$repro[[i]][j], digits)
  }
}

componentR2 <- c(SLA = 24.10, PC1 = 21.82, LT = 12.49, CNratio = 18.25,
                 PC2 = 6.06)
for (t in names(componentR2)) {
  addCheck("Component models", paste(t, "-> SIV R² (%)"), componentR2[[t]],
           100 * summary(modD[[t]])$r.squared, 2)
}
addCheck("Component models", "SIV -> LDMC R² (%)", 3.77,
         100 * summary(modS$LDMC)$r.squared, 2)

componentCoefs |>
  transmute(path,
            published = map_chr(pub, \(x)
              sprintf("%s [%s, %s]", x[1], x[2], x[3])),
            reproduced = map_chr(repro, \(x)
              sprintf("%.4f [%.4f, %.4f]", x[1], x[2], x[3]))) |>
  kable(caption = "Path coefficients reported in the Results.") |>
  kable_styling(full_width = FALSE)
```

# Structural equation models (Table 1, Fig. 5)

The historical model list has 54 SEMs, each fitted without and with plot type (108 in all), built from the component models above for six traits (SLA, LT, LDMC, C:N, PC1, PC2) and nine biological hypotheses (BH).

```{r sem-list}
# One SEM per trait for each hypothesis, in the historical order. Model
# components follow the historical psem() argument order.
buildSems <- function(p) {
  list(
    i    = map(traitNames, \(t) psem(modL_sivTrait[[p]][[t]], modD[[t]])),
    ii   = map(traitNames, \(t) psem(modL_sivTrait[[p]][[t]], modS[[t]])),
    vi   = map(traitNames, \(t) psem(modD[[t]], modL_siv[[p]])),
    vii  = map(traitNames, \(t) psem(modS[[t]], modL_trait[[p]][[t]])),
    iii  = map(traitNames, \(t)
      psem(modL_sivTrait[[p]][[t]], modS0[[t]], modD0)),
    iv   = map(traitNames, \(t) psem(modL_siv[[p]], modD0, modS0[[t]])),
    viii = map(traitNames, \(t) psem(modD[[t]], modS0[[t]], modL0[[p]])),
    ix   = map(traitNames, \(t) psem(modS[[t]], modD0, modL0[[p]])),
    v    = map(traitNames, \(t)
      psem(modL_trait[[p]][[t]], modS0[[t]], modD0)))
}
sems <- map(names(plotTypes), \(p) {
  s <- buildSems(p)
  tibble(plotType = p, BH = rep(names(s), each = length(traitNames)),
         trait = rep(traitNames, length(s)),
         sem = unname(list_flatten(s)))
}) |>
  list_rbind() |>
  mutate(id = row_number(), .before = 1)

# HISTORICAL BUG: model 55 (plot type, BH i, SLA) was written as
# psem(L8.1ct, S1c) instead of psem(L8.1ct, D1c), so it duplicates the
# BH ii model 61 but is labelled BH i.
idBug <- with(sems, id[plotType == "plot" & BH == "i" & trait == "SLA"])
sems$sem[[idBug]] <- psem(modL_sivTrait$plot$SLA, modS$SLA)
```

```{r sem-selection}
# AICc for every SEM; a model that fails becomes NA and drops silently
# (historical tryCatch)
sems$AICc <- map_dbl(sems$sem, \(m)
  tryCatch(AIC(m, aicc = TRUE), error = \(e) NA_real_))
sems <- sems |>
  mutate(dAICc = AICc - min(AICc, na.rm = TRUE),
         # Fig. 5 weights: from unrounded dAICc, relative likelihoods
         # rounded to 5 dp, weights to 3 dp (historical rounding)
         RL = round(exp(-0.5 * dAICc), 5),
         weight_sow = round(RL / sum(RL), 3)) |>
  arrange(dAICc) |>
  # Table 1 weights: from dAICc rounded to 2 dp (historical rounding)
  mutate(rank = row_number(),
         dAICc_tab = round(dAICc, 2),
         weight = round(round(exp(-0.5 * dAICc_tab), 5) /
                          sum(round(exp(-0.5 * dAICc_tab), 5),
                              na.rm = TRUE), 3))

# R² (marginal, %) per component and Fisher's C. The historical code read
# Fisher's C from summary(), which fails under piecewiseSEM 2.1.2 for
# SEMs with an intercept-only component. For those, dSep() + fisherC() is
# used instead; it can differ from summary() in the third decimal.
sems <- sems |>
  mutate(
    R2 = map_chr(sem, \(m) tryCatch(
      paste(round(100 * rsquared(m)$Marginal, 2), collapse = " / "),
      error = \(e) NA_character_)),
    FisherC_source = "summary",
    FisherC = map_dbl(sem, \(m) tryCatch(
      as.numeric(summary(m, .progressBar = FALSE)$Cstat[1]),
      error = \(e) NA_real_)),
    FisherC_source = if_else(is.na(FisherC), "fisherC", FisherC_source),
    FisherC = if_else(
      is.na(FisherC),
      map_dbl(sem, \(m) tryCatch(
        fisherC(dSep(m, .progressBar = FALSE))$Fisher.C,
        error = \(e) NA_real_)),
      FisherC),
    formula = map_chr(sem, \(m) {
      comps <- m[map_lgl(m, \(x) inherits(x, c("lm", "merMod")))]
      map_chr(comps, \(x) deparse1(formula(x))) |>
        str_remove(fixed(" + (1 | plot) + (logba | species)")) |>
        paste(collapse = "; ")
    }))

sumOfWeights <- sems |>
  summarise(SoW = sum(weight_sow), .by = BH) |>
  arrange(desc(SoW))
```

```{r table1-checks}
table1Pub <- tribble(
  ~rank, ~BH,   ~R2,             ~FisherC, ~AICc, ~dAICc, ~weight,
  1,     "vi",  "35.55 / 24.10", 0.383,    27.09, 0.00,   0.313,
  2,     "vi",  "35.55 / 21.82", 1.469,    28.41, 1.32,   0.162,
  3,     "i",   "34.86 / 24.10", 0.173,    29.78, 2.69,   0.082,
  4,     "vii", "33.21 / 3.77",  2.849,    30.08, 2.98,   0.071,
  5,     "vi",  "35.55 / 12.49", 3.027,    30.30, 3.20,   0.063,
  6,     "i",   "36.36 / 21.82", 0.627,    30.34, 3.25,   0.062,
  7,     "ii",  "35.37 / 3.77",  0.743,    30.49, 3.39,   0.057,
  8,     "vi",  "35.55 / 18.25", 3.651,    31.05, 3.96,   0.043,
  9,     "vi",  "35.55 / 6.06",  4.758,    32.39, 5.30,   0.022,
  10,    "i",   "35.98 / 12.49", 2.453,    32.59, 5.50,   0.020,
  25,    "iv",  "33.40 / 18.25", NA,       38.61, 11.51,  0.001)

table1Cmp <- table1Pub |>
  left_join(select(sems, rank, BH, R2, FisherC, AICc, dAICc, weight),
            by = "rank", suffix = c("_pub", ""))

for (i in seq_len(nrow(table1Cmp))) {
  r <- table1Cmp[i, ]
  lab <- paste("rank", r$rank)
  addCheck("Table 1", paste(lab, "AICc"), r$AICc_pub, r$AICc, 2)
  addCheck("Table 1", paste(lab, "dAICc"), r$dAICc_pub, r$dAICc, 2)
  addCheck("Table 1", paste(lab, "weight"), r$weight_pub, r$weight, 3)
  if (!is.na(r$FisherC_pub)) {
    addCheck("Table 1", paste(lab, "Fisher's C"), r$FisherC_pub,
             r$FisherC, 3)
  }
}
# R²: Table 1 lists the canopy model first, whatever the component order,
# so the values are compared as sets
table1R2Match <- with(table1Cmp,
  map2_lgl(R2_pub, R2, \(a, b) {
    pa <- sort(as.numeric(str_split_1(a, " / ")))
    pb <- sort(as.numeric(str_split_1(b, " / ")))
    identical(pa, pb)
  }))
bhMatch <- table1Cmp$BH_pub == table1Cmp$BH

sowPub <- c(vi = 61.3, i = 19.2, vii = 9.3, ii = 7.8)
for (h in names(sowPub)) {
  addCheck("Fig. 5", paste("summed weight (%), BH", h), sowPub[[h]],
           100 * sumOfWeights$SoW[sumOfWeights$BH == h], 1)
}
```

All `r sum(!is.na(sems$AICc))` of `r nrow(sems)` SEMs have an AICc. The ten top-ranked SEMs and rank 25, the first with plot type, match Table 1:

```{r table1}
sems |>
  filter(rank <= 10 | rank == 25) |>
  select(rank, BH, formula, R2, FisherC, AICc, dAICc = dAICc_tab, weight) |>
  kable(digits = 3,
        caption = "Table 1 reproduced (R² in %, per component model).") |>
  kable_styling(font_size = 12)
```

The R² columns match the published values for `r sum(table1R2Match)` of `r length(table1R2Match)` rows, and the BH labels for `r sum(bhMatch)` of `r length(bhMatch)`. `r if (!bhMatch[table1Cmp$rank == 25]) paste0("Both mismatches are at rank 25. Its structure (canopy ~ BA + C:N + plot type; C:N ~ SIV) is hypothesis ", table1Cmp$BH[table1Cmp$rank == 25], " in the code, but Table 1 labels it \"iv\", a transcription error in the paper; and its canopy-model R² is ", str_extract(table1Cmp$R2[table1Cmp$rank == 25], "[0-9.]+$"), " against the published 33.40, a rounding difference.")`

```{r fig5, fig.width = 6, fig.height = 4, fig.cap = "Figure 5 reproduced: summed Akaike weights per biological hypothesis."}
sumOfWeights |>
  mutate(BH = fct_reorder(BH, SoW)) |>
  ggplot(aes(SoW * 100, BH)) +
  geom_col(fill = "#40798C") +
  geom_text(aes(label = sprintf("%.1f%%", SoW * 100)), hjust = -0.1,
            size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Summed Akaike weight (%)", y = "Biological hypothesis") +
  theme_bw()
```

# Fig. 4

```{r fig4, fig.width = 7, fig.height = 11, fig.cap = "Figure 4 reproduced: a) PCA of leaf traits; b) SIV (square-rooted, as in the published figure) against SLA."}
pcaScores <- bind_cols(code = traitsOnly$code, as_tibble(leafPca$x))
pcaArrows <- as_tibble(leafPca$rotation, rownames = "trait") |>
  mutate(trait = recode(trait, CNratio = "C:N"))
fig4a <- ggplot(pcaScores, aes(PC1, PC2)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_text(aes(label = code)) +
  geom_segment(data = pcaArrows,
               aes(x = 0, y = 0, xend = 3 * PC1, yend = 3 * PC2),
               colour = "forestgreen", linewidth = 1,
               arrow = arrow(length = unit(0.2, "cm"))) +
  geom_text(data = pcaArrows, aes(4 * PC1, 3.5 * PC2, label = trait),
            colour = "forestgreen", size = 5) +
  labs(x = sprintf("PC 1 (%.1f%% of variance)", 100 * pcaVariance[["PC1"]]),
       y = sprintf("PC 2 (%.1f%% of variance)", 100 * pcaVariance[["PC2"]]),
       title = "a)") +
  theme_bw()
fig4b <- litter |>
  distinct(code, SLA, iv.wet) |>
  ggplot(aes(SLA * 100, iv.wet, label = code)) +
  geom_text() +
  labs(x = expression(SLA~(cm^2~g^-1)),
       y = "Habitat specialisation (sqrt of swamp indicator value)",
       title = "b)") +
  theme_bw()
patchwork::wrap_plots(fig4a, fig4b, ncol = 1)
```

# Appendix B: canopy production models (Table S3)

```{r table-s3}
library(MuMIn)
# Canopy production models without BA ("N" models), for Appendix B
fitLmerNoBa <- function(fixed) {
  f <- reformulate(c(if (is.null(fixed)) "1" else fixed,
                     "(1 | plot)", "(logba | species)"),
                   response = "loglitter")
  eval(bquote(lmer(.(f), data = litter)))
}
traitsS3 <- c(`1` = "SLA", `2` = "LT", `3` = "LDMC", `4` = "CNratio")
modsS3 <- c(
  list(L0 = modL0$noPlot, L0t = modL0$plot),
  imap(traitsS3, \(t, k)
    list(modL_trait$noPlot[[t]], modL_trait$plot[[t]]) |>
      set_names(paste0("L", k), paste0("L", k, "t"))) |>
    unname() |>
    list_flatten(),
  list(L7 = modL_siv$noPlot),
  set_names(modL_sivTrait$noPlot, paste0("L8.", 1:6, "c")),
  list(N0 = fitLmerNoBa(NULL), N0t = fitLmerNoBa("plotwet")),
  imap(traitsS3, \(t, k)
    list(fitLmerNoBa(t), fitLmerNoBa(c("plotwet", t))) |>
      set_names(paste0("N", k), paste0("N", k, "t"))) |>
    unname() |>
    list_flatten(),
  list(N7 = fitLmerNoBa("iv.wet")),
  map(set_names(traitNames, paste0("N8.", 1:6, "c")), \(t)
    fitLmerNoBa(c("iv.wet", t))))

modelSelS3 <- model.sel(modsS3)
tableS3 <- tibble(model = rownames(modelSelS3)) |>
  bind_cols(
    map(rownames(modelSelS3), \(m) {
      cm <- coef(summary(modsS3[[m]]))
      terms <- c("logba", "plotwet", "SLA", "LT", "LDMC", "CNratio",
                 "iv.wet", "PC1", "PC2")
      map_chr(set_names(terms), \(j)
        if (j %in% rownames(cm)) {
          sprintf("%.2f (±%.2f)", cm[j, "Estimate"], cm[j, "Std. Error"])
        } else NA_character_) |>
        as_tibble_row()
    }) |>
      list_rbind(),
    as.data.frame(modelSelS3)[c("df", "logLik", "AICc", "delta",
                                "weight")] |>
      mutate(across(everything(), as.numeric)) |>
      as_tibble()) |>
  mutate(R2adj = map_dbl(model, \(m) tryCatch(
    attr(r.squaredLR(modsS3[[m]]), "adj.r.squared"),
    error = \(e) NA_real_)))

tableS3 |>
  kable(digits = 3, caption = "Table S3 reproduced.") |>
  kable_styling(font_size = 11) |>
  scroll_box(height = "400px")
```

Like Table S1, the published Table S3 could not be checked, because the Supplementary Information was not available.

# Phenology GAMM (Fig. 3)

The GAMM relates BA-normalised leaf litter production per collection (non-liana species) to plot condition, a smooth of date and weather in the 30 days before each collection, using NEA daily records.

```{r weather}
# Monthly NEA files, as read by the historical code. HISTORICAL QUIRK: an
# offset vector of months stops the reading at June 2019, although
# collections run to 30 July 2019. Rain at Upper Peirce (S69) is replaced
# by Mandai (S40) for March-June 2019, when S69 lacks most of its data.
dirNea <- "../../TFERP/LTFEM_aux/data/nea"
monthsRead <- format(seq(as.Date("2018-06-01"), as.Date("2019-06-01"),
                         by = "month"), "%Y%m")
readNea <- function(folder, station, months) {
  map(months, \(m)
    read.csv(file.path(dirNea, folder,
                       sprintf("DAILYDATA_%s_%s.csv", station, m)),
             fileEncoding = "CP1252", check.names = FALSE,
             colClasses = "character")[, 1:11]) |>
    list_rbind() |>
    set_names(c("station", "year", "month", "day", "rain", "r30", "r60",
                "r120", "tmean", "tmax", "tmin")) |>
    mutate(date1 = make_date(as.integer(year), as.integer(month),
                             as.integer(day)))
}
mandaiMonths <- c("201903", "201904", "201905", "201906")
temperature <- readNea("S122_khatib", "S122", monthsRead)
rain <- bind_rows(
  readNea("S69_upper_peirce_reservoir", "S69",
          setdiff(monthsRead, mandaiMonths)),
  readNea("S40_mandai", "S40", mandaiMonths))

# Missing values are coded with an em dash. HISTORICAL: missing rain is
# set to 0; a missing daily mean temperature takes that month's mean; a
# missing max or min stays NA.
weather <- temperature |>
  select(date1, tmean, tmax, tmin) |>
  left_join(select(rain, date1, rain), by = "date1") |>
  mutate(rain = as.numeric(if_else(rain == "—", "0", rain)),
         across(c(tmean, tmax, tmin), \(x) suppressWarnings(as.numeric(x))),
         ym = format(date1, "%Y%m")) |>
  mutate(tmean = if_else(is.na(tmean), mean(tmean, na.rm = TRUE), tmean),
         .by = ym)
```

```{r gamm-data}
# For each collection date d: the 30 days before it (d-30 to d-1), and
# the 30 days before that (d-60 to d-31) for the lagged variables
windowMean <- function(x, d, from, to) {
  map_dbl(d, \(di) mean(x[weather$date1 >= di - from &
                            weather$date1 <= di - to]))
}
windowMax <- function(x, d) {
  map_dbl(d, \(di) max(x[weather$date1 >= di - 30 &
                           weather$date1 <= di - 1], na.rm = TRUE))
}
gammData <- standLitter |>
  mutate(rf = windowMean(weather$rain, date1, 30, 1),
         rf_lag = windowMean(weather$rain, date1, 60, 31),
         temp = windowMean(weather$tmean, date1, 30, 1),
         temp_lag = windowMean(weather$tmean, date1, 60, 31),
         temp_max = windowMax(weather$tmax, date1),
         temp_min = -windowMax(-weather$tmin, date1))

# Compare with the historical GAMM input file
gammHist <- read.csv("data/data for GAMM.csv") |>
  mutate(date1 = as.Date(date1))
gammCmp <- inner_join(gammData, gammHist, by = c("Plot", "date1"),
                      suffix = c("", "_hist"))
gammVars <- c("Dry.Mass", "ba", "prod", "rf", "rf_lag", "temp", "temp_lag",
              "temp_max", "temp_min")
gammMaxDiff <- map_dbl(set_names(gammVars), \(v)
  max(abs(gammCmp[[v]] - gammCmp[[paste0(v, "_hist")]])))
addCheck("GAMM data", "rows (data for GAMM.csv)", nrow(gammHist),
         nrow(gammData), 0)
addCheck("GAMM data", "rows matched", nrow(gammHist), nrow(gammCmp), 0)
addCheck("GAMM data", "max abs difference, all columns", 0,
         max(gammMaxDiff), 6)

# Collections whose 30-day window runs past the weather data
nTruncated <- sum(gammData$date1 - 1 > max(weather$date1))
```

The rebuilt GAMM data (`r nrow(gammData)` plot–collections) match the historical input file `data for GAMM.csv`, with a largest absolute difference of `r format(signif(max(gammMaxDiff), 2), scientific = TRUE)` across litter, BA and the six weather variables. Because the weather read stops on `r format(max(weather$date1), "%d %B %Y")`, `r nTruncated` collections in July 2019 have truncated 30-day windows; `audit/09_gamm_fullwindows.R` shows that complete windows leave the results materially unchanged.

```{r gamm-fit}
library(mgcv)
library(gamm4)
# The historical model selection, verbatim apart from the data name.
# CAUTION: the rows must be sorted by Plot (the random-effect grouping), as
# the historical merge() left them. Otherwise gamm4 0.2-6 returns the same
# estimates but wrong standard errors for the parametric terms.
tll <- gammData |>
  arrange(Plot, date1) |>
  as.data.frame()
uGammFull <- uGamm(log(prod) ~ Condition + s(as.numeric(date1))
  + rf + rf_lag + temp + temp_lag
  + temp_max + temp_min
  + Condition:rf + Condition:rf_lag + Condition:temp + Condition:temp_lag
  + Condition:temp_max + Condition:temp_min,
  random = ~(1|Plot), data = tll, lme4 = TRUE)
dredged <- dredge(uGammFull, m.lim = c(0, 4), rank = "AICc",
  subset = !((rf && rf_lag) || (temp && temp_lag) ||
    (temp_max && temp_min) || (temp && temp_max) ||
    (temp && temp_min) || (temp_lag && temp_max) || (temp_lag && temp_min)))

best <- gamm4(log(prod) ~ s(as.numeric(date1)) + Condition + temp,
              random = ~(1|Plot), data = tll)
bestCoefs <- summary(best$gam)$p.table
waldCi <- function(term) {
  bestCoefs[term, "Estimate"] +
    c(est = 0, lwr = -1.96, upr = 1.96) * bestCoefs[term, "Std. Error"]
}
# Dry relative to wet (the reference level is dry, D)
dryWet <- set_names(exp(-waldCi("ConditionW"))[c("est", "upr", "lwr")],
                    c("est", "lwr", "upr"))
tempEffect <- waldCi("temp")
r2Top <- r.squaredLR(get.models(dredged, subset = 1)[[1]])

addCheck("GAMM", "top-model weight > 0.90", 1,
         as.numeric(dredged$weight[1] > 0.9), 0)
addCheck("GAMM", "dAICc of 2nd model", 5.47, dredged$delta[2], 2)
addCheck("GAMM", "adj R² (LR) of top model (%)", 75,
         100 * attr(r2Top, "adj.r.squared"), 0)
addCheck("GAMM", "dry/wet multiplier", 2.01, dryWet[[1]], 2)
addCheck("GAMM", "dry/wet lower 95%", 1.57, dryWet[[2]], 2)
addCheck("GAMM", "dry/wet upper 95%", 2.57, dryWet[[3]], 2)
addCheck("GAMM", "temperature estimate", 0.40, tempEffect[["est"]], 2)
addCheck("GAMM", "temperature lower 95%", 0.27, tempEffect[["lwr"]], 2)
addCheck("GAMM", "temperature upper 95%", 0.53, tempEffect[["upr"]], 2)
```

The top-ranked GAMM includes plot condition and mean temperature (Akaike weight `r round(dredged$weight[1], 3)`; the next model is `r round(dredged$delta[2], 2)` AICc units behind), with adjusted likelihood-ratio *R*^2^ = `r round(100 * attr(r2Top, "adj.r.squared"))`%. BA-normalised production was `r round(dryWet[[1]], 2)` [`r round(dryWet[[2]], 2)`, `r round(dryWet[[3]], 2)`] times higher in dry than in wet plots, and rose by `r sprintf("%.2f", tempEffect[["est"]])` [`r round(tempEffect[["lwr"]], 2)`, `r round(tempEffect[["upr"]], 2)`] on the log scale per °C. The effect sizes differ from the published ones in the second decimal place. No confidence-interval method tried (Wald, *t*, profile) recovers them exactly; package-version drift since 2021 is the likely cause (unverified).

```{r fig3, fig.width = 8, fig.height = 9, fig.cap = "Figure 3 reproduced: a) BA-normalised leaf litter production with GAMM predictions (± 1.96 SE) for wet and dry plots; b) mean daily temperature over the 30 days before each collection."}
newDates <- sort(unique(tll$date1))
gammPred <- map(c(W = "W", D = "D"), \(cond) {
  nd <- tibble(date1 = newDates, Condition = cond,
               temp = tll$temp[match(newDates, tll$date1)])
  p <- predict(best$gam, newdata = nd, se.fit = TRUE)
  mutate(nd, fit = p$fit, se = p$se.fit)
}) |>
  list_rbind() |>
  mutate(Condition = recode(Condition, W = "Wet", D = "Dry"))

fig3a <- ggplot(gammPred, aes(date1, fit, colour = Condition,
                              fill = Condition)) +
  geom_ribbon(aes(ymin = fit - 1.96 * se, ymax = fit + 1.96 * se),
              colour = NA, alpha = 0.3) +
  geom_line(aes(linetype = Condition), linewidth = 1.2) +
  geom_point(data = mutate(tll, Condition = recode(Condition, W = "Wet",
                                                   D = "Dry")),
             aes(y = log(prod), shape = Condition), alpha = 0.7) +
  scale_y_continuous(breaks = log(c(4, 8, 16, 32, 64, 128)),
                     labels = c(4, 8, 16, 32, 64, 128)) +
  scale_colour_manual(values = c(Wet = "#1F363D", Dry = "#8D8266")) +
  scale_fill_manual(values = c(Wet = "#40798C", Dry = "#9EC1A3")) +
  labs(x = NULL, title = "a)", colour = "Plot condition",
       fill = "Plot condition", linetype = "Plot condition",
       shape = "Plot condition",
       y = expression(BA*"-normalised litter production"~
                        (g~fortnight^-1~m^-2~m^-2))) +
  theme_bw()
fig3b <- tll |>
  distinct(date1, temp) |>
  ggplot(aes(date1, temp)) +
  geom_line(linewidth = 1) +
  labs(x = "Date", y = "Mean daily temperature (°C)", title = "b)") +
  theme_bw()
patchwork::wrap_plots(fig3a, fig3b, ncol = 1, heights = c(2, 1))
```

# Summary of checks

```{r summary-checks}
checksSummary <- checks |>
  summarise(checks = n(), matched = sum(match), .by = section)
```

Of `r nrow(checks)` comparisons with published or historical values, `r sum(checks$match)` match at the published precision. The mismatches are:

```{r mismatches}
checks |>
  filter(!match) |>
  kable(caption = "Comparisons that do not match.") |>
  kable_styling(full_width = FALSE)
```

```{r all-checks}
checksSummary |>
  kable(caption = "Checks by section.") |>
  kable_styling(full_width = FALSE)

checks |>
  kable(caption = "All checks.") |>
  kable_styling(full_width = FALSE, font_size = 11) |>
  scroll_box(height = "400px")
```

# Reference outputs

```{r reference-outputs}
# Written for comparison with corrected_reanalysis.Rmd
dir.create("reference", showWarnings = FALSE)
write_csv(select(litterSppPlot, -start), "reference/historical_species_plot.csv")
write_csv(bind_cols(traitsOnly, as_tibble(leafPca$x)),
          "reference/historical_species_traits_pca.csv")
write_csv(as_tibble(leafPca$rotation, rownames = "trait"),
          "reference/historical_pca_loadings.csv")
write_csv(
  bind_rows(
    imap(modD, \(m, t) tibble(model = paste(t, "-> SIV"),
                              term = names(coef(m)),
                              estimate = unname(coef(m)),
                              r2 = summary(m)$r.squared)) |>
      list_rbind(),
    imap(modS, \(m, t) tibble(model = paste("SIV ->", t),
                              term = names(coef(m)),
                              estimate = unname(coef(m)),
                              r2 = summary(m)$r.squared)) |>
      list_rbind(),
    tibble(model = "SIV -> canopy (L7)",
           term = names(fixef(modL_siv$noPlot)),
           estimate = unname(fixef(modL_siv$noPlot)), r2 = NA)),
  "reference/historical_component_models.csv")
write_csv(select(sems, -sem), "reference/historical_sem_table.csv")
write_csv(sumOfWeights, "reference/historical_sum_of_weights.csv")
write_csv(tableS1, "reference/historical_tableS1.csv")
write_csv(tableS3, "reference/historical_tableS3.csv")
write_csv(
  tibble(quantity = c("weight_top", "dAICc_2nd", "r2adj_top",
                      "dry_wet", "dry_wet_lwr", "dry_wet_upr",
                      "temp", "temp_lwr", "temp_upr"),
         value = c(dredged$weight[1], dredged$delta[2],
                   attr(r2Top, "adj.r.squared"), dryWet, tempEffect)),
  "reference/historical_gamm.csv")
write_csv(checks, "reference/historical_checks.csv")
```

The main outputs are in `reference/`: `r paste(list.files("reference"), collapse = ", ")`.

# Session information

```{r session-info}
sessionInfo()
```
